### 2024-02-02 products/manuscript/Manuscript copy.qmd
project part 1

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Xueyan HU
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: dataanalysis-references.bib
csl: apa.csl
editor: 
  markdown: 
    wrap: sentence
---

The structure below is one possible setup for a data analysis project (including the course project).
For a manuscript, adjust as needed.
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format.
[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
You can switch to other formats, like html or pdf.
See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

In this project, I will be concentrating on analyzing my own data gotten from the experiment that I carried on by my own.
And I will be trying to figure out if there will be any relation between interested variables and observations.

{{< pagebreak >}}

# Introduction

## General Background Information

The core idea of this research comes from the following facts.
First, sprouts are usually considered nutritious but they are consumed in raw mainly for foods like salad and sandwiches.
Therefore, there have been a lot of foodborne pathogen outbreaks occurred across the states in the decades [@miyahira2021], which means that there is an urgent need to reduce the contamination of sprouts with foodborne pathogens.

At the meantime, a recently study has revealed that a newly derived chemical could enhence the immune resistance of some certain plants to the pathogens including bacteria and virus but not by killing them directly[@manosalva2015].

Based on the facts above, a hypothesis is raised that if this chemical can also be applied on seedling seeds to control the level of contamination during sprouting process.

## Description of data and data source

The data is obtained from experiment that I did it by myself, which has 160 observations and several variables.
The data is the result of bacterial populations of foodborne pathogens on sprouts that are collected from different types of plants at different germinating time points under different treatments for the seeds

## Questions/Hypotheses to be addressed

I would like to figure out the effect of different treatments on controlling bacterial populations as well as the factors also may be related to the effectiveness, including seed type for sprouts, bacterial strain type infecting sprouts, etc.
So the treatment type will be the most important predictor that I will focus on.
And I am willing to explore if there is any treatment that could reduce bacterial population level on seedling sprouts.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Schematic of workflow

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
Many more style files for almost any journal [are available](https://www.zotero.org/styles).
You also specify the location of your bibtex reference file in the YAML.
You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2024-02-23 R/analysis-code/exploratory-analysis.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

linear model applied

```{r}
# Fit the ANOVA model
model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)

# Summarize the ANOVA results
linear_model <- summary(model)

# Save this table

anovatable_file = here("results", "tables", "anovatable.rds")
saveRDS(linear_model, file = anovatable_file)
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.

### 2024-02-23 R/eda-code/eda-v2.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

--- 2024-02-23 R/analysis-code/exploratory-analysis.qmd
@@
 ---
-title: "An example exploratory analysis script"
-author: "Andreas Handel"
-date: "10/28/2021"
+title: "Project data cleaning"
+author: "Xueyan Hu"
+date: "2023-02-22"
 output: html_document
@@
 
-I'm only showing it the way where the code is included in the file. 
-As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.
-
-But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)
-
-Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
-And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.
+This is essentially the same as the other `exploratory_analysis` Quarto file, but now the code is not inside this file.
+Instead, it is pulled in from the R script `exploratorycode.R` using the code chunk labels.
 
@@
 # Setup
+Load the chunk.
 
-```{r}
-#load needed packages. make sure they are installed.
-library(here) #for data loading/saving
-library(dplyr)
-library(skimr)
-library(ggplot2)
+```{r, include=FALSE, cache=FALSE}
+knitr::read_chunk('edacode.R')
+```
+
+Load the packages.
+```{r,packages, echo=FALSE,message=FALSE}
 ```
@@
 
-```{r}
-#Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
-#load data
-mydata <- readRDS(data_location)
+```{r,loaddata}
 ```
@@
 
-```{r}
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+```{r,table1}
+```
+
+We are saving the results to the `results` folder.
+Depending on how many tables/figures you have, it might make sense to have separate folders for each.
+And/or you could have separate folders for exploratory tables/figures and for final tables/figures.
+Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.
+
+# Data exploration through figures
+
+Seed type first: bar chart
+
+```{r,seed}
+```
+
+Then goes strain type: box plot
+
+```{r,strain}
+```
+
+Now the effect of the treatment: bar chart with color gradient as an indicator
+
+```{r,treatment}
+```
+
+The growth trend of the bacterial population over time: scatterplot with line
+
+```{r,day}
 ```
@@
 
-```{r}
-# Fit the ANOVA model
-model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)
-
-# Summarize the ANOVA results
-linear_model <- summary(model)
-
-# Save this table
-
-anovatable_file = here("results", "tables", "anovatable.rds")
-saveRDS(linear_model, file = anovatable_file)
+```{r,modelfit}
 ```

### 2024-02-23 R/eda-code/eda.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

--- 2024-02-23 R/analysis-code/exploratory-analysis.qmd
@@
 
-linear model applied
+We are saving the results to the `results` folder.
+Depending on how many tables/figures you have, it might make sense to have separate folders for each.
+And/or you could have separate folders for exploratory tables/figures and for final tables/figures.
+Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.
+
+# Data exploration through figures
+
+Seed type first: bar chart
 
 ```{r}
-# Fit the ANOVA model
-model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)
+# Calculate average population by seed type
+mydata_seed <- mydata %>%
+  group_by(Seed) %>%
+  summarize(mean_Selective = mean(Selective),
+            sd_Selective = sd(Selective))
 
-# Summarize the ANOVA results
-linear_model <- summary(model)
+# Create the bar chart with standard deviation lines
+bp1 <- ggplot(mydata_seed, aes(x = Seed, y = mean_Selective, fill = Seed)) +
+  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
+  geom_errorbar(data = mydata_seed, aes(x = Seed, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "black"),
+                position = position_dodge(width = 0.1), width = 0.05) +
+  labs(title = "The population difference by seed type",
+       x = "Seed Type",
+       y = "Average Population") +
+  scale_color_manual(values = c("Alfalfa" = "lightblue", "Fenugreek" = "pink")) +
+  theme_minimal()
+plot(bp1)
+figure_file = here("results","figures", "seed_type.png")
+ggsave(filename = figure_file, plot=bp1) 
+```
 
-# Save this table
+Then goes strain type: box plot
 
-anovatable_file = here("results", "tables", "anovatable.rds")
-saveRDS(linear_model, file = anovatable_file)
+```{r}
+# Create boxplot for bacterial population by 
+bp2 <- ggplot(mydata, aes(x = Strain, y = Selective)) +
+  geom_boxplot() +
+  labs(title = "The population difference by treatment",
+       x = "Strain type",
+       y = "Average Population") +
+  theme_minimal()
+plot(bp2)
+figure_file = here("results","figures", "strain_type.png")
+ggsave(filename = figure_file, plot=bp2) 
+```
+
+Now the effect of the treatment: bar chart with color gradient as an indicator
+
+```{r}
+# Calculate average population by seed type
+mydata_treatment <- mydata %>%
+  group_by(Treatment) %>%
+  summarize(mean_Selective = mean(Selective),
+            sd_Selective = sd(Selective))
+
+# Create a continuous variable for fill
+mydata_treatment$fill_color <- scales::rescale(mydata_treatment$mean_Selective)
+
+# Create the bar chart with standard deviation lines
+bp3 <- ggplot(mydata_treatment, aes(x = Treatment, y = mean_Selective, fill = fill_color)) +
+  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
+  geom_errorbar(data = mydata_treatment, aes(x = Treatment, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "grey"),
+                position = position_dodge(width = 0.1), width = 0.05) +
+  labs(title = "The population difference by treatment",
+       x = "Treatment",
+       y = "Average Population") +
+  scale_fill_gradient(low = "yellow", high = "red") +  # Adjust the colors as needed
+  scale_color_manual(values = c("grey")) +  # Use grey color for error bar color
+  guides(color = "none") +  # Turn off both fill and color legends
+  theme_minimal()
+plot(bp3)
+figure_file = here("results","figures", "treatment.png")
+ggsave(filename = figure_file, plot=bp3) 
+```
+
+The growth trend of the bacterial population over time: scatterplot with line
+
+```{r}
+# Calculate the average population for each day
+mydata_day <- aggregate(Selective ~ Day, data = mydata, FUN = mean)
+
+# Draw a point plot
+sp4 <- ggplot(mydata_day, aes(x = Day, y = Selective)) +
+  geom_point(size = 2, color = "blue") +
+  geom_line(aes(group = 1), color = "red") +
+  labs(title = "The growth trend of the bacterial population over time",
+       x = "Day",
+       y = "Average Bacterial Population") +
+  theme_minimal()
+plot(sp4)
+figure_file = here("results","figures", "timepoint.png")
+ggsave(filename = figure_file, plot=sp4) 
+```
 ```

### 2024-02-23 R/processing-code/processingfile-v1.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

---
title: "Project data cleaning"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

Note that for functions that come from specific packages (instead of base R), I often specify both package and function like so:
package::function() that's not required one could just call the function specifying the package makes it clearer where the function "lives",
but it adds typing. You can do it either way.

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","raw_data.xlsx")
rawdata <- readxl::read_excel(data_location)
```

# Check data

First we can look at the codebook

```{r}
codebook <- readxl::read_excel(data_location, sheet ="Codebook")
print(codebook)
```

Several ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

# Cleaning

The data looks generally clear before cleaning. So I just want to make sure that "Day" could be defined as a categorical variable but not a numeric variable.

```{r}
data1 <- rawdata %>%
  mutate(Day = paste0("Day", substring(Day, 2)))
skimr::skim(data1)
head(data1)
```

# Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data:
http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
processeddata <- data1
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2024-02-23 R/processing-code/processingfile-v2.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

--- 2024-02-23 R/processing-code/processingfile-v1.qmd
@@
 
-# Processing script
+# Processing script with code pulled in
 
-This contains the same code and comments/information as `processingcode.R`.
-
-This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.
-
-See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.
+This is essentially the same as the other Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script using the code chunk labels.
 
@@
 
-Load needed packages. make sure they are installed.
+This needs to run to load the R script so we can include the code chunks below.
 
-```{r}
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
-library(here) #to set paths
+Note that you can pull in more than one R script if you want.
+
+```{r, include=FALSE, cache=FALSE}
+knitr::read_chunk('processingcode.R')
+```
+
+Load needed packages. Instead of having R commands here, this just pulls in the code from the R script we loaded above and labelled.
+
+```{r, packages, message = FALSE, warning = FALSE}
 ```
@@
 
-```{r}
-# path to data
-# note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw-data","raw_data.xlsx")
-rawdata <- readxl::read_excel(data_location)
+```{r, loaddata}
 ```
 
-# Check data
-
-First we can look at the codebook
-
-```{r}
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-print(codebook)
-```
+# Explore data
 
@@
 
-```{r}
-dplyr::glimpse(rawdata)
-summary(rawdata)
-head(rawdata)
-skimr::skim(rawdata)
+```{r, exploredata}
 ```
@@
 
-```{r}
-data1 <- rawdata %>%
-  mutate(Day = paste0("Day", substring(Day, 2)))
-skimr::skim(data1)
-head(data1)
+```{r, cleandata1}
 ```
@@
 # Save data
+
+All done, data is clean now. Time to save.
+Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.
 
@@
 
-```{r}
-processeddata <- data1
-save_data_location <- here::here("data","processed-data","processeddata.rds")
-saveRDS(processeddata, file = save_data_location)
+```{r, savedata}
 ```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g.
-keep individuals with some missing information).

### 2024-02-23 products/manuscript/Manuscript copy.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
+title: "Examine the efficacy of ascaroside #18 in controling Escherichia Coli on alfalfa and fenugreek seeds and sprouts"
 subtitle: ""
-author: Xueyan HU
+author: Xueyan Hu
 date: "`r Sys.Date()`"
@@
 
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
-
 {{< pagebreak >}}
@@
 
-*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
+Since the dataset is quite well-organized and no any value missing, the cleaning step is relatively easy.
+I only changed the category names.
+
+My main interest is to test if the chemical treatment is effective or not compared to the control treated groups.
+Also, anova model will be applied to check which variable will make significant different to the bacterial population.
+So boxplot, bar chart and line chart will be created according to the analysis direction.
 
@@
 
-*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*
+All of the data in this dataset was collected by me in 2022 after experiment was conducted.
 
 ## Data import and cleaning
+
+The detailed raw data and processed data are stored in 'data' folder.
 

### 2024-02-23 R/analysis-code/exploratorycode.R

@@ -7,7 +7,7 @@ library(ggplot2)
 
 ## ---- loaddata --------
 #Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
+data_location <- here::here("starter-analysis-exercise","data","processed-data","processeddata.rds")
 #load data
 mydata <- readRDS(data_location)
 
@@ -15,30 +15,30 @@ mydata <- readRDS(data_location)
 summary_df = skimr::skim(mydata)
 print(summary_df)
 # save to file
-summarytable_file = here("results", "summarytable.rds")
+summarytable_file = here("starter-analysis-exercise","results", "summarytable.rds")
 saveRDS(summary_df, file = summarytable_file)
 
 ## ---- height --------
 p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
 plot(p1)
-figure_file = here("results","height_distribution.png")
+figure_file = here("starter-analysis-exercise","results","height_distribution.png")
 ggsave(filename = figure_file, plot=p1) 
 
 ## ---- weight --------
 p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
 plot(p2)
-figure_file = here("results","weight_distribution.png")
+figure_file = here("starter-analysis-exercise","results","weight_distribution.png")
 ggsave(filename = figure_file, plot=p2) 
 
 ## ---- fitfig1 --------
 p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
 plot(p3)
-figure_file = here("results","height_weight.png")
+figure_file = here("starter-analysis-exercise","results","height_weight.png")
 
 ## ---- fitfig2 --------
 p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
 plot(p4)
-figure_file = here("results","height_weight_stratified.png")
+figure_file = here("starter-analysis-exercise","results","height_weight_stratified.png")
 ggsave(filename = figure_file, plot=p4) 
 
 

### 2024-02-23 R/analysis-code/statistical-analysis.R

@@ -22,35 +22,24 @@ mydata <- readRDS(data_location)
 ######################################
 
 ############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
-
-lmfit1 <- lm(Height ~ Weight, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
-
-#look at fit results
-print(lmtable1)
-
-# save fit results table  
-table_file1 = here("results", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
+#### summary table
+summary_df = skimr::skim(mydata)
+print(summary_df)
+# save to file
+summarytable_file = here("results", "tables", "summarytable.rds")
+saveRDS(summary_df, file = summarytable_file)
 
 ############################
-#### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
+#### model fit
+# Fit the ANOVA model
+model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)
 
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
+# Summarize the ANOVA results
+linear_model <- summary(model)
 
-# place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
+# Save this table
 
-#look at fit results
-print(lmtable2)
-
-# save fit results table  
-table_file2 = here("results", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
+anovatable_file = here("results", "tables", "anovatable.rds")
+saveRDS(linear_model, file = anovatable_file)
 
   

### 2024-02-23 R/eda-code/edacode.R

@@ -7,7 +7,7 @@ library(ggplot2)
 
 ## ---- loaddata --------
 #Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
+data_location <- here::here("data","processed-data","processeddata.rds")
 #load data
 mydata <- readRDS(data_location)
 
@@ -15,30 +15,92 @@ mydata <- readRDS(data_location)
 summary_df = skimr::skim(mydata)
 print(summary_df)
 # save to file
-summarytable_file = here("results", "summarytable.rds")
+summarytable_file = here("results", "tables", "summarytable.rds")
 saveRDS(summary_df, file = summarytable_file)
 
-## ---- height --------
-p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
-plot(p1)
-figure_file = here("results","height_distribution.png")
-ggsave(filename = figure_file, plot=p1) 
+## ---- seed --------
+# Calculate average population by seed type
+mydata_seed <- mydata %>%
+  group_by(Seed) %>%
+  summarize(mean_Selective = mean(Selective),
+            sd_Selective = sd(Selective))
 
-## ---- weight --------
-p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
-plot(p2)
-figure_file = here("results","weight_distribution.png")
-ggsave(filename = figure_file, plot=p2) 
+# Create the bar chart with standard deviation lines
+bp1 <- ggplot(mydata_seed, aes(x = Seed, y = mean_Selective, fill = Seed)) +
+  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
+  geom_errorbar(data = mydata_seed, aes(x = Seed, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "black"),
+                position = position_dodge(width = 0.1), width = 0.05) +
+  labs(title = "The population difference by seed type",
+       x = "Seed Type",
+       y = "Average Population") +
+  scale_color_manual(values = c("Alfalfa" = "lightblue", "Fenugreek" = "pink")) +
+  theme_minimal()
+plot(bp1)
+figure_file = here("results","figures", "seed_type.png")
+ggsave(filename = figure_file, plot=bp1) 
 
-## ---- fitfig1 --------
-p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-plot(p3)
[…]

### 2024-02-23 R/processing-code/processingcode.R

@@ -28,6 +28,7 @@ data_location <- here::here("data","raw_data","exampledata.xlsx")
 #package::function() that's not required one could just call the function
 #specifying the package makes it clearer where the function "lives",
 #but it adds typing. You can do it either way.
+data_location <- here::here("data","raw-data","raw_data.xlsx")
 rawdata <- readxl::read_excel(data_location)
 
 # We can look in the Codebook (second tab) for a variable explanation
@@ -52,68 +53,20 @@ skimr::skim(rawdata)
 
 
 ## ---- cleandata1 --------
-# Inspecting the data, we find some problems that need addressing:
-# First, there is an entry for height which says "sixty" instead of a number. 
-# Does that mean it should be a numeric 60? It somehow doesn't make
-# sense since the weight is 60kg, which can't happen for a 60cm person (a baby)
-# Since we don't know how to fix this, we might decide to remove the person.
-# This "sixty" entry also turned all Height entries into characters instead of numeric.
-# That conversion to character also means that our summary function isn't very meaningful.
-# So let's fix that first.
-
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
+# The data looks generally clear before cleaning. 
+#So I just want to make sure that "Day" could be 
+#defined as a categorical variable but not a numeric variable.
 
 
-# look at partially fixed data again
-skimr::skim(d1)
-hist(d1$Height)
+data1 <- rawdata %>%
+  mutate(Day = paste0("Day", substring(Day, 2)))
+skimr::skim(data1)
+head(data1)
 
 
-## ---- cleandata2 --------
-# Now we see that there is one person with a height of 6. 
-# that could be a typo, or someone mistakenly entered their height in feet.
-# If we don't know, we might need to remove this person.
-# But let's assume that we somehow know that this is meant to be 6 feet, so we can convert to centimeters.
-d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
-
-
-#height values seem ok now
-skimr::skim(d2)
-
-
-## ---- cleandata3 --------
-# now let's look at weight
-# there is a person with weight of 7000, which is impossible,
-# and one person with missing weight.
-# Note that the original data had an empty cell. 
-# The codebook says that's not allowed, it should have been NA.
-# R automatically converts empty values to NA.
[…]

### 2024-02-24 R/analysis-code/exploratory-analysis-v2.qmd
mistakes fixed

--- 2024-02-23 R/eda-code/eda-v2.qmd
@@
 ---
-title: "Project data cleaning"
-author: "Xueyan Hu"
-date: "2023-02-22"
+title: "An example exploratory analysis script with code pulled in"
+author: "Yao Lu, Andreas Handel"
+date: "2023-03-13"
 output: html_document
@@
 
-Also note that while here I split cleaning and exploring, this is iterative.
-You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.
-In general, as you explore, you'll find things that need cleaning.
-As you clean, you can explore more.
-Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file.
-Or split things in any other logical way.
-
-As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-Figures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data.
-
-Start by exploring one variable at a time.
-Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-
-Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.
-
 # Setup
@@
 ```{r, include=FALSE, cache=FALSE}
-knitr::read_chunk('edacode.R')
+knitr::read_chunk('exploratorycode.R')
 ```
@@
 
-```{r,table1}
+```{r,table 1}
 ```
@@
 
-# Data exploration through figures
+# Data exploration through through fit model
 
-Seed type first: bar chart
-
-```{r,seed}
+```{r,ANOVA table}
 ```
 
-Then goes strain type: box plot
+# LSD test
 
-```{r,strain}
+```{r,LSD test}
 ```
-
-Now the effect of the treatment: bar chart with color gradient as an indicator
-
-```{r,treatment}
-```
-
-The growth trend of the bacterial population over time: scatterplot with line
-
-```{r,day}
-```
-
-linear model applied
-
-```{r,modelfit}
-```
-
 # Notes

### 2024-02-24 R/analysis-code/exploratory-analysis.qmd
mistakes fixed

@@
 saveRDS(linear_model, file = anovatable_file)
+
+# Fit an new ANOVA model
+model1 <- aov(Selective ~ Treatment, data = mydata)
+
+# Perform LSD test
+lsd_result <- LSD.test(model1, "Treatment", alpha = 0.05)
+
+# Display the results
+print(lsd_result)
+
+# save LSD result
+
+lsdtable_file = here("results", "tables", "lsdtable.rds")
+saveRDS(lsd_result, file = lsdtable_file)
 ```

### 2024-02-24 products/manuscript/Manuscript copy.qmd
mistakes fixed

@@
 ---
-
-The structure below is one possible setup for a data analysis project (including the course project).
-For a manuscript, adjust as needed.
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
 
@@
 
-## Data aquisition
+## Data acquisition
 
@@
 The detailed raw data and processed data are stored in 'data' folder.
+And the code for data cleaning is in the 'processing-code' subfolder under 'R' folder.
+Since the name of the categories in variable "Day" is a bit confusing, I changed it to the form of day+time point0,1,3,5,7.
 
-*Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along.*
+And the reason why I only choose the colony count from selective media is that supportive media may have more colonies that are not specific target strain due to its ingredient feature.
+So the count from selective media could be more reliable.
 
@@
 
-*Explain anything related to your statistical analyses.*
+By fitting the data with ANOVA model, the result will show whether each independent variable has main effect to the dependent variable through the P value.
+And the summary table is stored in 'table' subfolder under 'result' folder.
+
+Then for the efficacy of chemical, a LSD test is applied and the significant level will be presented for the 4 treatments.
+
+Bar chart, box plot will show the difference of bacterial populations between 2 types of seeds, 2 types of strains, and among 4 treatments.
+As well as line graph for the bacteria growth trend during sprouting process.
 
@@
 
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
+My main interest here is the realtionship between treatment and bacterial populations.
+Based on the ANOVA table, seed type, treatment and day are the variables that have significant effect on the dependent variables; while strain type is not.
+And there is no significant difference between 2 replications, which means the experiment mayvbe replicable.
+Furthermore, I ran a LSD test to check the significant level for each 4 treatment and got the result as follows.
+
+Table 1.
+The result of ANOVA model.
+
+| Variable  | Df  | Sum Sq | Mean Sq | F value | Pr(F)    |
+|-----------|-----|--------|---------|---------|----------|
+| Seed      | 1   | 95.8   | 95.81   | 43.287  | 7.53e-10 |
+| Strain    | 1   | 0.1    | 0.10    | 0.047   | 0.829    |
+| Treatment | 3   | 322.3  | 107.43  | 48.538  | \< 2e-16 |
+| Day       | 4   | 113.8  | 28.46   | 12.859  | 5.12e-09 |
+| Rep       | 1   | 0.4    | 0.42    | 0.190   | 0.664    |
+| Residuals | 149 | 329.8  | 2.21    |         |          |
+
+Table 2.
+Treatment summary table.
+
+| Treatment type     | Mean population | Significant level |
+|--------------------|:---------------:|-------------------|
+| Control 2          |      4.79       | a                 |
+| Control 1          |      4.49       | a                 |
+| Chemical treated 2 |      3.23       | b                 |
+| Chemical treated 1 |      1.18       | c                 |
+
+The figures created are for some basic concepts about the population difference among the categories.
+
+![](images/seed_type.png){fig-align="center" width="500"}
+
+Figure 1.
+The population difference between alfalfa and fenugreek seeds.
+
+![](images/strain_type.png){fig-align="center" width="500"}
+
+Figure 2.
+The population difference between F strain and G strain (no significant difference).
+
+![](images/treatment.png){fig-align="center" width="500"}
+
+Figure 3.
+The effect of the chemical treatment application representing by color.
+
+![](images/timepoint.png){fig-align="center" width="500" height="450"}
+
+Figure 4.
+The growth trend of bacteria population during sprouting process.
 

### 2024-02-24 R/analysis-code/exploratorycode.R

@@ -18,27 +18,24 @@ print(summary_df)
 summarytable_file = here("starter-analysis-exercise","results", "summarytable.rds")
 saveRDS(summary_df, file = summarytable_file)
 
-## ---- height --------
-p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
-plot(p1)
-figure_file = here("starter-analysis-exercise","results","height_distribution.png")
-ggsave(filename = figure_file, plot=p1) 
 
-## ---- weight --------
+## ---- ANOVA table --------
 p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
 plot(p2)
 figure_file = here("starter-analysis-exercise","results","weight_distribution.png")
-ggsave(filename = figure_file, plot=p2) 
-
-## ---- fitfig1 --------
-p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-plot(p3)
-figure_file = here("starter-analysis-exercise","results","height_weight.png")
-
-## ---- fitfig2 --------
-p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
-plot(p4)
-figure_file = here("starter-analysis-exercise","results","height_weight_stratified.png")
-ggsave(filename = figure_file, plot=p4) 
+ggsave(filename = figure_file, plot=p2)
 
 
+## ---- LSD test --------
+# Perform LSD test
+lsd_result <- LSD.test(model1, "Treatment", alpha = 0.05)
+
+# Display the results
+print(lsd_result)
+
+# save LSD result
+
+lsdtable_file = here("results", "tables", "lsdtable.rds")
+saveRDS(lsd_result, file = lsdtable_file)
+
+ 

### 2024-02-24 R/analysis-code/statistical-analysis.R

@@ -1,45 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-############################
-#### summary table
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
-
-############################
-#### model fit
-# Fit the ANOVA model
-model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)
-
-# Summarize the ANOVA results
-linear_model <- summary(model)
-
-# Save this table
-
-anovatable_file = here("results", "tables", "anovatable.rds")
-saveRDS(linear_model, file = anovatable_file)
-
-  

### 2024-03-19 R/eda-code/eda.qmd
change dataset 1

@@
 ---
-title: "Project data cleaning"
+title: "Project eda"
 author: "Xueyan Hu"
@@
 library(ggplot2)
+library(maps)
+library(gt)
 ```
@@
 #Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
+data_location <- here::here("data","processed-data","processeddata_cancer.rds")
 #load data
@@
 
-# Data exploration through tables
-
-Showing a bit of code to produce and save a summary table.
+# bar chart for crude prevalence according to states
 
 ```{r}
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+# Group by state and calculate the total population and cancer cases
+state_total <- mydata %>%
+  group_by(StateDesc) %>%
+  summarize(TotalPopulation = sum(PopulationCount),
+            TotalCancerCases = sum(Cancer_Case))
+
+# Calculate the prevalence for each state
+state_total <- state_total %>%
+  mutate(TotalPrevalence = TotalCancerCases / TotalPopulation)
+
+# View the resulting dataframe
+print(state_total)
+
+# Load the map data for the US states
+us_states <- map_data("state")
+print(us_states)
+
+# Convert StateDesc column to lowercase
+state_total$StateDesc <- tolower(state_total$StateDesc)
+
+# Merge the map data with the state_totals dataframe
+map_data <- merge(us_states, state_total, by.x = "region", by.y = "StateDesc", all.x = TRUE)
+
+# Plot the map
+map <- ggplot() +
+  geom_map(data = map_data, map = map_data,
+           aes(x = long, y = lat, map_id = region, fill = TotalPrevalence),
+           color = "black", size = 0.1) +
+  scale_fill_gradient(name = "TotalPrevalence", low = "lightblue", high = "darkblue") +
+  labs(title = "Cancer Prevalence by State") +
+  theme_void() +
+  theme(legend.position = "right")
+plot(map)
+```
+
+# List 10 cities that has the highest cancer prevalence
+
+```{r}
+top_10_highest <- mydata %>%
+  top_n(10, Data_Value) %>%
+  select(Year, StateAbbr, CityName, Data_Value)
+
+print(top_10_highest)
+
+# create a table
+table <- top_10_highest %>%
+  gt() %>%
+  tab_spanner(
+    label = md('**Year**'),
+    columns = c('Year')
+  ) %>%
+  tab_spanner(
+    label = md('**State**'),
+    columns = c('StateAbbr')
+  ) %>%
+  tab_spanner(
+    label =  md('**Cancer Prevalence**'),
+    columns = c('Data_Value')
+  ) %>%
+  tab_header(
+    title = '2017 top 10 cities which have highest cancer rates'
+  ) %>%
+  opt_stylize(style = 6, color = 'gray')
 ```
@@
 
-# Data exploration through figures
-
-Seed type first: bar chart
+# fit a simple linear model
 
 ```{r}
-# Calculate average population by seed type
-mydata_seed <- mydata %>%
-  group_by(Seed) %>%
-  summarize(mean_Selective = mean(Selective),
-            sd_Selective = sd(Selective))
-
-# Create the bar chart with standard deviation lines
-bp1 <- ggplot(mydata_seed, aes(x = Seed, y = mean_Selective, fill = Seed)) +
-  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
-  geom_errorbar(data = mydata_seed, aes(x = Seed, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "black"),
-                position = position_dodge(width = 0.1), width = 0.05) +
-  labs(title = "The population difference by seed type",
-       x = "Seed Type",
-       y = "Average Population") +
-  scale_color_manual(values = c("Alfalfa" = "lightblue", "Fenugreek" = "pink")) +
+# create boxplot of Fare by Passenger class
+bp1 <- ggplot(mydata, aes(x = as.factor(Pclass), y = Fare)) +
+  geom_boxplot() +
+  labs(title = "Boxplot of Ticket Fare by Passenger Class",
+       x = "Passenger Class",
+       y = "Fare") +
   theme_minimal()

### 2024-03-19 R/processing-code/processingfile-v1.qmd
change dataset 1

@@
 
-Load needed packages. make sure they are installed.
+### load packages
 
@@
 
-# Data loading
-
-Note that for functions that come from specific packages (instead of base R), I often specify both package and function like so:
-package::function() that's not required one could just call the function specifying the package makes it clearer where the function "lives",
-but it adds typing. You can do it either way.
+### loading data
 
 ```{r}
-# path to data
-# note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw-data","raw_data.xlsx")
-rawdata <- readxl::read_excel(data_location)
-```
-
-# Check data
-
-First we can look at the codebook
-
-```{r}
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-print(codebook)
+data_location <- here::here("data","raw-data","500_Cities__Cancer__excluding_skin_cancer__among_adults_aged___18_years_20240314.csv")
+rawdata <- read.csv(data_location)
 ```
@@
 Several ways of looking at the data
+
+The meaning of each variable can be checked [online](https://www.kaggle.com/datasets/yasserh/titanic-dataset/data).
 
@@
 
-# Cleaning
+### data cleaning
 
-The data looks generally clear before cleaning. So I just want to make sure that "Day" could be defined as a categorical variable but not a numeric variable.
+First, only crude prevalence is interested since we don't know the age structure here.
+And the prevalence of each city is the summary of all the census tracts from this city, so City will be selected only for further analysis.
+There are also 2 summed data including all the US prevalence that will not be used for analysis.
 
 ```{r}
-data1 <- rawdata %>%
-  mutate(Day = paste0("Day", substring(Day, 2)))
-skimr::skim(data1)
-head(data1)
+d1 <- rawdata %>%
+  filter(GeographicLevel == "City", DataValueTypeID == "CrdPrv", StateAbbr != "US")
 ```
 
-# Save data
-
-Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
-This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data:
-http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
+Dropping uninterested variables for a smaller dataset.
 
 ```{r}
-processeddata <- data1
-save_data_location <- here::here("data","processed-data","processeddata.rds")
+d2 <- d1 %>%
+  select(Year, StateAbbr, StateDesc, CityName, DataValueTypeID, Data_Value,PopulationCount)
+```
+
+Generating a new variable in case it will be needed later.
+
+```{r}
+d3 <- d2 %>%
+  mutate(Cancer_Case = round((Data_Value / 100) * PopulationCount))
+```
+
+Re-check with the data
+
+```{r}
+dplyr::glimpse(d3)
+summary(d3)
+head(d3)
+skimr::skim(d3)
+```
+
+### Save data
+
+Finally, we save the clean data as RDS file.
+I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
+This preserves coding like factors, characters, numeric, etc.
+If you save as CSV, that information would get lost.
+However, CSV is better for sharing with others since it's plain text.
+If you do CSV, you might want to write down somewhere what each variable is.
+
+See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
+
+```{r}
+processeddata <- d3
+save_data_location <- here::here("data","processed-data","processeddata_cancer.rds")
 saveRDS(processeddata, file = save_data_location)

### 2024-03-19 R/processing-code/processingcode.R

@@ -18,9 +18,6 @@ library(skimr) #for nice visualization of data
 library(here) #to set paths
 
 ## ---- loaddata --------
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
 
 #load data. 
 #note that for functions that come from specific packages (instead of base R)
@@ -28,12 +25,8 @@ data_location <- here::here("data","raw_data","exampledata.xlsx")
 #package::function() that's not required one could just call the function
 #specifying the package makes it clearer where the function "lives",
 #but it adds typing. You can do it either way.
-data_location <- here::here("data","raw-data","raw_data.xlsx")
-rawdata <- readxl::read_excel(data_location)
-
-# We can look in the Codebook (second tab) for a variable explanation
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-print(codebook)
+data_location <- here::here("data","raw-data","500_Cities__Cancer__excluding_skin_cancer__among_adults_aged___18_years_20240314.csv")
+rawdata <- read.csv(data_location)
 
 
 ## ---- exploredata --------
@@ -52,31 +45,38 @@ skimr::skim(rawdata)
 
 
 
-## ---- cleandata1 --------
+## ---- cleandata --------
 # The data looks generally clear before cleaning. 
 #So I just want to make sure that "Day" could be 
 #defined as a categorical variable but not a numeric variable.
 
 
-data1 <- rawdata %>%
-  mutate(Day = paste0("Day", substring(Day, 2)))
-skimr::skim(data1)
-head(data1)
+# choose needed categories
+d1 <- rawdata %>%
+  filter(GeographicLevel == "City", DataValueTypeID == "CrdPrv", StateAbbr != "US")
 
+# drop uninterested variables for a smaller dataset
+d2 <- d1 %>%
+  select(Year, StateAbbr, StateDesc, CityName, DataValueTypeID, Data_Value,PopulationCount)
 
-# look at partially if it is changed or not
-skimr::skim(data1)
-head(data1)
+# generating a new variable
+d3 <- d2 %>%
+  mutate(Cancer_Case = round((Data_Value / 100) * PopulationCount))
 
-
+# re-check with the data
+dplyr::glimpse(d3)
+summary(d3)
+head(d3)
+skimr::skim(d3)
 
 ## ---- savedata --------
 # all done, data is clean now. 
 # Let's assign at the end to some final variable
 # makes it easier to add steps above
-processeddata <- data1
[…]

### 2024-03-19 R/eda-code/eda.qmd
eda map

--- 2024-02-23 R/analysis-code/exploratory-analysis.qmd
@@
 ---
-title: "An example exploratory analysis script"
-author: "Andreas Handel"
-date: "10/28/2021"
+title: "Project eda"
+author: "Xueyan Hu"
+date: "2023-02-22"
 output: html_document
@@
 
-# Data exploration through tables
-
-Showing a bit of code to produce and save a summary table.
+# bar chart for crude prevalence according to states
 
 ```{r}
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+# Group by state and calculate the total population and cancer cases
+state_total <- mydata %>%
+  group_by(StateDesc) %>%
+  summarize(TotalPopulation = sum(PopulationCount),
+            TotalCancerCases = sum(Cancer_Case))
+
+# Calculate the prevalence for each state
+state_total <- state_total %>%
+  mutate(TotalPrevalence = TotalCancerCases / TotalPopulation)
+
+# View the resulting dataframe
+print(state_total)
+
+# Load usmap package and get the state data
+library(usmap)
+
+usmap_state_data <- us_map(regions = "states")
+
+# There are a few state names that are cut off in state_total, but the column
+# is already alphabetical so we can replace it.
+# To use plot_usmap() we have to have a column named `state` or `fips`.
+state_total$state <- usmap_state_data$full
+print(state_total)
+
+# Now we can try to make a usmap plot
+plot_usmap("states", data = state_total, values = "TotalPrevalence") +
+  scale_fill_viridis_c()
 ```
 
-linear model applied
+# List 10 cities that has the highest cancer prevalence
 
 ```{r}
-# Fit the ANOVA model
-model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)
+top_10_highest <- mydata %>%
+  top_n(10, Data_Value) %>%
+  select(Year, StateAbbr, CityName, Data_Value)
 
-# Summarize the ANOVA results
-linear_model <- summary(model)
+print(top_10_highest)
 
-# Save this table
+# create a table
+table <- top_10_highest %>%
+  gt() %>%
+  tab_spanner(
+    label = md('**Year**'),
+    columns = c('Year')
+  ) %>%
+  tab_spanner(
+    label = md('**State**'),
+    columns = c('StateAbbr')
+  ) %>%
+  tab_spanner(
+    label =  md('**Cancer Prevalence**'),
+    columns = c('Data_Value')
+  ) %>%
+  tab_header(
+    title = '2017 top 10 cities which have highest cancer rates'
+  ) %>%
+  opt_stylize(style = 6, color = 'gray')
+```
 
-anovatable_file = here("results", "tables", "anovatable.rds")
-saveRDS(linear_model, file = anovatable_file)
+We are saving the results to the `results` folder.
+Depending on how many tables/figures you have, it might make sense to have separate folders for each.
+And/or you could have separate folders for exploratory tables/figures and for final tables/figures.
+Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.
+
+# fit a simple linear model
+
+Before we started analysis, we had an assumption
+
+```{r}
+# Fit linear regression model
+lm1 <- lm(Data_Value ~ PopulationCount, data = mydata)
+
+# Print summary of the model
+summary(lm1)
 ```
+So the hypothesis is not supported by the linear model fitting result. There is no obvious evidence that the cancer prevalence in larger cities (higher population)
 

### 2024-03-23 R/analysis-code/exploratory-analysis.qmd
change dataset and finish part 3

--- 2024-02-23 R/analysis-code/exploratory-analysis.qmd
@@
 title: "An example exploratory analysis script"
-author: "Andreas Handel"
-date: "10/28/2021"
+author: "Xueyan Hu"
+date: "03/22/2024"
 output: html_document
@@
 
-# Data exploration through tables
+# fit simple linear model
 
-Showing a bit of code to produce and save a summary table.
+First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.
 
 ```{r}
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+# Fit multinomial logistic regression model
+#library(nnet)
+#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)
+
+# Summary of the model
+#summary(multinom_model)
+```
+I will try to figure it out later how I could make it fitting.
+
+Then, I would like to see how drinking water plus physical activity affect obesity.
+
+```{r}
+# linear model of BMI by water and FAF
+lm1 <- lm(BMI ~ Water + FAF + Water * FAF, data = mydata)
+lm1_result <- summary(lm1)
+
+# save the result
+lm_file = here("results", "tables", "lm1table.rds")
+saveRDS(lm1_result, file = lm_file)
 ```
 
-linear model applied
+In addition, I want to fit model for alcohol consumption and obesity level.
 
 ```{r}
-# Fit the ANOVA model
-model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)
+mydata$Alcohol <- as.factor(mydata$Alcohol)
+lm2 <- lm(BMI ~ Alcohol, data = mydata)
+lm2_result <- summary(lm2)
 
-# Summarize the ANOVA results
-linear_model <- summary(model)
+# save the result
+lm_file = here("results", "tables", "lm2table.rds")
+saveRDS(lm2_result, file = lm_file)
+```
+I think it makes a little bit sense to me that there is a connection effect of water and physical activity on BMI, which might be supportive for what we learn that drinking more water and doing more exercise will lower the risk of obesity.
 
-# Save this table
+Last, I will try to use all predictors that I believe may affect BMI to fit a model.
 
-anovatable_file = here("results", "tables", "anovatable.rds")
-saveRDS(linear_model, file = anovatable_file)
+```{r}
+mydata1 <- mydata %>%
+  select(History, Water, Alcohol, FAF, MTRANS, BMI)
+lm3 <- lm(BMI ~ ., data = mydata1)
+lm3_result <- summary(lm3)
+
+# save the result
+lm_file = here("results", "tables", "lm3table.rds")
+saveRDS(lm3_result, file = lm_file)
 ```
-
 # Notes

### 2024-03-23 R/eda-code/eda-v2.qmd
change dataset and finish part 3

@@
 
-# Data exploration through tables
+# Data exploration through figures
 
-Showing a bit of code to produce and save a summary table.
+Figures of population distribution by gender and age.
 
-```{r,table1}
+```{r,p1 to p3}
 ```
@@
 
+Relationship between obesity level and gender/age.
+
+```{r,p4 and p5}
+```
+
+Violin plot of obesity level by family history.
+
+```{r,p6}
+```
+
 # Data exploration through figures
 
-Seed type first: bar chart
+Summary tables.
 
-```{r,seed}
+```{r,table}
 ```
 
-Then goes strain type: box plot
-
-```{r,strain}
-```
-
-Now the effect of the treatment: bar chart with color gradient as an indicator
-
-```{r,treatment}
-```
-
-The growth trend of the bacterial population over time: scatterplot with line
-
-```{r,day}
-```
-
-linear model applied
-
-```{r,modelfit}
-```
-
-# Notes
+# note
 

### 2024-03-23 R/eda-code/eda.qmd
change dataset and finish part 3

@@
 library(ggplot2)
-library(maps)
-library(gt)
 ```
@@
 #Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata_cancer.rds")
+data_location <- here::here("data","processed-data","processeddata.rds")
 #load data
@@
 
-# bar chart for crude prevalence according to states
+# gender and age distribution for the cohort under investigation
 
 ```{r}
-# Group by state and calculate the total population and cancer cases
-state_total <- mydata %>%
-  group_by(StateDesc) %>%
-  summarize(TotalPopulation = sum(PopulationCount),
-            TotalCancerCases = sum(Cancer_Case))
+# Convert numeric values to labels
+mydata$Gender <- ifelse(mydata$Gender == 1, "Male", "Female")
 
-# Calculate the prevalence for each state
-state_total <- state_total %>%
-  mutate(TotalPrevalence = TotalCancerCases / TotalPopulation)
+# bar chart of gender distribution
+bc1 <- ggplot(mydata, aes(x = Gender)) +
+  geom_bar(fill = c("pink", "skyblue"), color = "black", width = 0.5) +
+  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
+  labs(x = "Gender", y = "Count", title = "Bar Chart of Gender Distribution") +
+  theme_minimal()
+plot(bc1)
 
-# View the resulting dataframe
-print(state_total)
-
-# Load usmap package and get the state data
-library(usmap)
-
-usmap_state_data <- us_map(regions = "states")
-
-# There are a few state names that are cut off in state_total, but the column
-# is already alphabetical so we can replace it.
-# To use plot_usmap() we have to have a column named `state` or `fips`.
-state_total$state <- usmap_state_data$full
-print(state_total)
-
-# Now we can try to make a usmap plot
-plot_usmap("states", data = state_total, values = "TotalPrevalence") +
-  scale_fill_viridis_c()
+# histogram of population by age
+hp2 <- ggplot(mydata, aes(x = Age)) +
+  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
+  labs(x = "Age", y = "Count", title = "Histogram of Age distribution") +
+  stat_bin(binwidth = 5, geom = "text", aes(label = ..count..),
+           vjust = -0.5, color = "black", size = 3)  +
+  theme_minimal()
+plot(hp2)
+ # bar chart of obesity level distribution
+bc3 <- ggplot(mydata, aes(x = Obesity)) +
+  geom_bar() +
+  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
+  scale_x_discrete(labels = c("Under", "Normal", "OW1", "OW2", "Obese1", "Obese2", "Obese3")) +
+  labs(x = "Obesity level", y = "Count", title = "Bar Chart of obesity Distribution") +
+  theme_minimal()
+plot(bc3)
 ```
 
-# List 10 cities that has the highest cancer prevalence
+# Explore a little bit deeper for Gender and Age
 
 ```{r}
-top_10_highest <- mydata %>%
-  top_n(10, Data_Value) %>%
-  select(Year, StateAbbr, CityName, Data_Value)
+unique(mydata$Obesity)
+# obesity level by gender
+sbp4 <- ggplot(mydata, aes(x = Obesity, fill = Gender)) +
+  geom_bar(position = "stack") +
+  labs(x = "Obesity Level", y = "Count", title = "Gender Distribution by Obesity Level") +
+  scale_fill_manual(values = c("skyblue", "pink"), labels = c("Male", "Female")) +
+  scale_x_discrete(labels = c("Under", "Normal", "OW1", "OW2", "Obese1", "Obese2", "Obese3")) +
+  theme_minimal()
+plot(sbp4)
 
-print(top_10_highest)
+# Obesity level by age
+# Categorize age into groups
+mydata <- mydata %>%
+  mutate(Age_Group = cut(Age, breaks = c(0, 20, 30, 40, 50, 60, 70, Inf),
+                         labels = c("0-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71+")))
 
-# create a table
-table <- top_10_highest %>%
-  gt() %>%
-  tab_spanner(
-    label = md('**Year**'),
-    columns = c('Year')
-  ) %>%
-  tab_spanner(
-    label = md('**State**'),
-    columns = c('StateAbbr')
-  ) %>%
-  tab_spanner(
-    label =  md('**Cancer Prevalence**'),
-    columns = c('Data_Value')
-  ) %>%
-  tab_header(
-    title = '2017 top 10 cities which have highest cancer rates'
-  ) %>%
-  opt_stylize(style = 6, color = 'gray')
+sbp5 <- ggplot(mydata, aes(x = Obesity, fill = Age_Group)) +
+  geom_bar(position = "stack",stat = "count") +
+  coord_flip() + # Rotate the plot to make it horizontal
+  labs(x = "Obesity level", y = "Count", title = "Stacked Bar Chart of Obesity Level by Age") +
+  theme_minimal()
+plot(sbp5)
 ```
+I think the population distribution for age is not balanced since 20-30 age groups contains most of the observations so I will not explore the relationship between obesity and age.
+
+# The relationship between family obesity history and obesity level
+
+```{r}
+vp6 <-ggplot(mydata, aes(x = History, y = BMI, fill = History)) +
+  geom_violin() +  # Add violin geometry
+  geom_boxplot(width = 0.1, fill = "white", color = "black") +  # Add boxplot for better visualization
+  geom_jitter(aes(color = History), width = 0.2, alpha = 0.5, size = 1) +   # Add jittered scatterplot
+  labs(x = "Family History of Obesity", y = "BMI", title = "Violin Plot of BMI by History of Obesity") +
+  scale_fill_manual(values = c("yes" = "lightyellow", "no" = "lightgreen")) +  # Define colors
+  theme_minimal()
+plot(vp6)
+```
+# summary table
+
+```{r}
+summary1 <- summary(mydata)
+summary2 <-skim(mydata)
+summarytable_file = here("results", "tables", "summarytable1.rds")
+saveRDS(summary1, file = summarytable_file)
+summarytable_file = here("results", "tables", "summarytable2.rds")
+saveRDS(summary2, file = summarytable_file)
+```
+# Saving all of the figures
 
@@
 
-# fit a simple linear model
-
-Before we started analysis, we had an assumption
-
 ```{r}
-# Fit linear regression model
-lm1 <- lm(Data_Value ~ PopulationCount, data = mydata)
-
-# Print summary of the model
-summary(lm1)
+figure_file = here("results","figures","gender-distribution.png")
+ggsave(filename = figure_file, plot=bc1) 
+figure_file = here("results","figures","obesity-distribution.png")
+ggsave(filename = figure_file, plot=bc3) 
+figure_file = here("results","figures","obesity level by geder.png")
+ggsave(filename = figure_file, plot=sbp4) 
+figure_file = here("results","figures","obesity level by family history.png")
+ggsave(filename = figure_file, plot=vp6) 
 ```
-So the hypothesis is not supported by the linear model fitting result. There is no obvious evidence that the cancer prevalence in larger cities (higher population)
 

### 2024-03-23 R/processing-code/processingfile-v1.qmd
change dataset and finish part 3

@@
 ```{r}
-data_location <- here::here("data","raw-data","500_Cities__Cancer__excluding_skin_cancer__among_adults_aged___18_years_20240314.csv")
+data_location <- here::here("data","raw-data","ObesityDataSet_raw_and_data_sinthetic.csv")
 rawdata <- read.csv(data_location)
@@
 
+#### check on codebook
+
+```{r}
+codebook <- here::here("data","raw-data","codebook.xlsx")
+read_excel(codebook)
+```
 Several ways of looking at the data
@@
 
-First, only crude prevalence is interested since we don't know the age structure here.
-And the prevalence of each city is the summary of all the census tracts from this city, so City will be selected only for further analysis.
-There are also 2 summed data including all the US prevalence that will not be used for analysis.
+First generating a new variable BMI using height and weight, as well as change some variable names and categorical levels.
 
@@
 d1 <- rawdata %>%
-  filter(GeographicLevel == "City", DataValueTypeID == "CrdPrv", StateAbbr != "US")
+  mutate(BMI = Weight / (Height)^2) %>%
+  rename(History = family_history_with_overweight, Water = CH2O, Alcohol = CALC, Obesity = NObeyesdad) %>%
+  mutate(Gender = if_else(Gender == "Male", 1, 2))
+
+d2 <- d1 %>% 
+  mutate(Alcohol = case_when(
+    Alcohol == "no" ~ 1,
+    Alcohol == "Sometimes" ~ 2,
+    Alcohol == "Frequently" ~ 3,
+    Alcohol == "Always" ~ 4,
+    TRUE ~ NA_integer_  # Handle any other cases
+  ))
+# Print the updated dataframe
+head(d2)
 ```
 
-Dropping uninterested variables for a smaller dataset.
-
-```{r}
-d2 <- d1 %>%
-  select(Year, StateAbbr, StateDesc, CityName, DataValueTypeID, Data_Value,PopulationCount)
-```
-
-Generating a new variable in case it will be needed later.
+My interest variables are gender, age, alcohol consumption, water drinking, transportation, physical activity, and outcomes will be obesity level and BMI.
 
@@
 d3 <- d2 %>%
-  mutate(Cancer_Case = round((Data_Value / 100) * PopulationCount))
-```
-
-Re-check with the data
-
-```{r}
-dplyr::glimpse(d3)
-summary(d3)
+  select(Gender, Age, History, Water, Alcohol, FAF, MTRANS,BMI, Obesity)
 head(d3)
-skimr::skim(d3)
 ```
@@
 processeddata <- d3
-save_data_location <- here::here("data","processed-data","processeddata_cancer.rds")
+save_data_location <- here::here("data","processed-data","processeddata.rds")
 saveRDS(processeddata, file = save_data_location)

### 2024-03-23 products/manuscript/Manuscript copy.qmd
change dataset and finish part 3

--- 2024-02-23 products/manuscript/Manuscript copy.qmd
@@
 ---
-title: "Examine the efficacy of ascaroside #18 in controling Escherichia Coli on alfalfa and fenugreek seeds and sprouts"
+title: "Explore the factors that may be associated with obesity"
 subtitle: ""
@@
 
-The structure below is one possible setup for a data analysis project (including the course project).
-For a manuscript, adjust as needed.
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
 ```{r, echo=FALSE, message=FALSE}
@@
 
-In this project, I will be concentrating on analyzing my own data gotten from the experiment that I carried on by my own.
-And I will be trying to figure out if there will be any relation between interested variables and observations.
+In this project, I have found......
 
@@
 
-The core idea of this research comes from the following facts.
-First, sprouts are usually considered nutritious but they are consumed in raw mainly for foods like salad and sandwiches.
-Therefore, there have been a lot of foodborne pathogen outbreaks occurred across the states in the decades [@miyahira2021], which means that there is an urgent need to reduce the contamination of sprouts with foodborne pathogens.
+[Obesity](https://www.who.int/news-room/fact-sheets/detail/obesity-and-overweight)![]() is a chronic complex disease defined by excessive fat deposits that can impair health.
+Obesity can lead to increased risk of type 2 diabetes and heart disease, it can affect bone health and reproduction, it increases the risk of certain cancers.
 
-At the meantime, a recently study has revealed that a newly derived chemical could enhence the immune resistance of some certain plants to the pathogens including bacteria and virus but not by killing them directly[@manosalva2015].
+In the recent decades, obesity has become a more and more common and severe problem across the world.
 
-Based on the facts above, a hypothesis is raised that if this chemical can also be applied on seedling seeds to control the level of contamination during sprouting process.
+Worldwide adult obesity has more than doubled since 1990, and adolescent obesity has quadrupled.
+In 2022, 1 in 8 people in the world were living with obesity.
+
+There is no doubt that obesity is a multifactorial disease due to obesogenic environments, psycho-social factors and genetic variants in most cases, but personal behavior could also has correlation with the result of overweight and obesity.
+
+There is a lot of discussion on the Internet and social media about fitness and weight loss, and people assume that physical activity can lose weight as a fact.
+Also KOLs ususally encourage people to drink more water for higher metabolic levels.
+But I have not seen a clear data to prove it.
+At the same time, the effect of physical activity to lose weight is also different from person to person.
+
+In addition, according to what I learned from nutrition course, alcohol consumption may be related to weight gain because [alcohol metabolism](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6527027/) and carbohydrate metabolism both go through the same pathway/cycle in the middle and late stages.
 
@@
 
-The data is obtained from experiment that I did it by myself, which has 160 observations and several variables.
-The data is the result of bacterial populations of foodborne pathogens on sprouts that are collected from different types of plants at different germinating time points under different treatments for the seeds
+The data is obtained from a website [UCI Machine Learning Repository](https://archive.ics.uci.edu/dataset/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition) which have collected the survey results of individuals from the countries of Mexico, Peru and Colombia on their eating habits and physical condition, as well as their estimation of obesity levels.
 
@@
 
-I would like to figure out the effect of different treatments on controlling bacterial populations as well as the factors also may be related to the effectiveness, including seed type for sprouts, bacterial strain type infecting sprouts, etc.
-So the treatment type will be the most important predictor that I will focus on.
-And I am willing to explore if there is any treatment that could reduce bacterial population level on seedling sprouts.
+I would like to figure out if higher frequency of physical activity and drinking more water could keep people away from obesity and whether heredity is an important factor for obesity.
+Also, I will focus on the relationship between alcohol consumption frequency and obesity level.
 
@@
 
-Since the dataset is quite well-organized and no any value missing, the cleaning step is relatively easy.
-I only changed the category names.
+Data cleaning including generating necessary variable, renaming categories and variables and so on.
 
-My main interest is to test if the chemical treatment is effective or not compared to the control treated groups.
-Also, anova model will be applied to check which variable will make significant different to the bacterial population.
-So boxplot, bar chart and line chart will be created according to the analysis direction.
+First to take a deeper look of the data, I would like to do some descriptive analysis including bar chart, boxplot, scatterplot etc. to indicate the distribution of the data.
+
+Then simple linear fit will be applied to explore the relationship between variables and outcomes (both categorical and numeric).
+
+Prediction will be made.
 
@@
 
-## Data aquisition
+## Data acquisition
 
-All of the data in this dataset was collected by me in 2022 after experiment was conducted.
+The dataset presents the results of a survey on the topic of obesity including personal physical conditions and answers to behavioral questions.
 
@@
 The detailed raw data and processed data are stored in 'data' folder.
+And the code for data cleaning is in the 'processing-code' subfolder under 'R' folder.
 
-*Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along.*
+Obesity level is a categorical variable defined by BMI range, so BMI as a new variable is created and it could be analyzed as a numeric variable.
+Also, to make the variable name easy to understand, I changed some abbreviated names to ones with fully spelled key words.
+Male is replaced by 1 and female by 2 under Gender variable for more convenient further analysis.
 
@@
 
-*Explain anything related to your statistical analyses.*
+Bar chart, box plot will show the distribution of obesity level by different variables, as well as the population features of the people under investigation.
+All figures will be saved in 'figure' subfolder under 'result' folder and shown in the manuscript.
+
+By fitting the data with simple linear model, the result will show whether each independent variable has main effect to the dependent variable through the P value.
+And the summary table is stored in 'table' subfolder under 'result' folder.
 
@@
 
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
+Bar chart shows that the genders of the cohort under investigation are balanced.
+
+Figure 1.
+Gender distribution.
+
+![](images/gender-distribution.png)
+
+It can be seen that almost two thirds of the people are identified with over weight.
+
+Figure 2.
+Obesity level distribution.
+
+![](images/obesity-distribution.png)
+
+There is no obvious pattern of obesity levels by gender.
+
+Figure 3.
+Gender distribution in each obesity levels.
+
+![](images/obesity%20level%20by%20gender.png)
+
+It is apparent that people with family obesity history may tend to be over weight according to figure 4 below.
+
+Figure 4.
+Violin plot of BMI by family obesity history.
+
+![](images/obesity level by family history-02.png)
 
@@
 
-*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
+Since some obvious relationship could be shown in figures, I choose other variables as predictors in exploratory analysis
+
+First, I would like to fit linear model for BMI by both water consumption and physical activity.
+And the result indicates that there is a connection effect between them two, which means that both of them contribution to obesity control.
+
+Also, I use single indicator, alcohol consumption and BMI as an outcome to fit linear model.
+
+In the last model I select all the variables that I think may have effect on BMI as predictors and BMI as outcome.
 
@@
 
+**Haven't updated yet but will do it later!!!**
+
 *What are the main take-home messages?*

### 2024-03-23 R/analysis-code/exploratorycode.R

@@ -11,31 +11,41 @@ data_location <- here::here("starter-analysis-exercise","data","processed-data",
 #load data
 mydata <- readRDS(data_location)
 
-## ---- table1 --------
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("starter-analysis-exercise","results", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+## ---- logistic --------
+# Fit multinomial logistic regression model
+#library(nnet)
+#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)
+
+# Summary of the model
+#summary(multinom_model)
 
 
-## ---- ANOVA table --------
-p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
-plot(p2)
-figure_file = here("starter-analysis-exercise","results","weight_distribution.png")
-ggsave(filename = figure_file, plot=p2)
+## ---- lm1 --------
+# linear model of BMI by water and FAF
+lm1 <- lm(BMI ~ Water + FAF + Water * FAF, data = mydata)
+lm1_result <- summary(lm1)
+
+# save the result
+lm_file = here("results", "tables", "lm1table.rds")
+saveRDS(lm1_result, file = lm_file)
 
 
-## ---- LSD test --------
-# Perform LSD test
-lsd_result <- LSD.test(model1, "Treatment", alpha = 0.05)
+## ---- lm2 --------
+mydata$Alcohol <- as.factor(mydata$Alcohol)
+lm2 <- lm(BMI ~ Alcohol, data = mydata)
+lm2_result <- summary(lm2)
 
-# Display the results
-print(lsd_result)
+# save the result
+lm_file = here("results", "tables", "lm2table.rds")
+saveRDS(lm2_result, file = lm_file)
 
-# save LSD result
+## ---- lm3 --------
+mydata1 <- mydata %>%
+  select(History, Water, Alcohol, FAF, MTRANS, BMI)
+lm3 <- lm(BMI ~ ., data = mydata1)
+lm3_result <- summary(lm3)
 
-lsdtable_file = here("results", "tables", "lsdtable.rds")
-saveRDS(lsd_result, file = lsdtable_file)
-
+# save the result
+lm_file = here("results", "tables", "lm3table.rds")
+saveRDS(lm3_result, file = lm_file)
  

### 2024-03-23 R/eda-code/edacode.R

@@ -11,96 +11,73 @@ data_location <- here::here("data","processed-data","processeddata.rds")
 #load data
 mydata <- readRDS(data_location)
 
-## ---- table1 --------
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results", "tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
+## ---- p1 to p3 --------
+# Convert numeric values to labels
+mydata$Gender <- ifelse(mydata$Gender == 1, "Male", "Female")
 
-## ---- seed --------
-# Calculate average population by seed type
-mydata_seed <- mydata %>%
-  group_by(Seed) %>%
-  summarize(mean_Selective = mean(Selective),
-            sd_Selective = sd(Selective))
-
-# Create the bar chart with standard deviation lines
-bp1 <- ggplot(mydata_seed, aes(x = Seed, y = mean_Selective, fill = Seed)) +
-  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
-  geom_errorbar(data = mydata_seed, aes(x = Seed, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "black"),
-                position = position_dodge(width = 0.1), width = 0.05) +
-  labs(title = "The population difference by seed type",
-       x = "Seed Type",
-       y = "Average Population") +
-  scale_color_manual(values = c("Alfalfa" = "lightblue", "Fenugreek" = "pink")) +
+# bar chart of gender distribution
+bc1 <- ggplot(mydata, aes(x = Gender)) +
+  geom_bar(fill = c("pink", "skyblue"), color = "black", width = 0.5) +
+  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
+  labs(x = "Gender", y = "Count", title = "Bar Chart of Gender Distribution") +
   theme_minimal()
-plot(bp1)
-figure_file = here("results","figures", "seed_type.png")
-ggsave(filename = figure_file, plot=bp1) 
+plot(bc1)
 
-## ---- strain --------
-# Create boxplot for bacterial population by 
-bp2 <- ggplot(mydata, aes(x = Strain, y = Selective)) +
-  geom_boxplot() +
-  labs(title = "The population difference by treatment",
-       x = "Strain type",
-       y = "Average Population") +
[…]

### 2024-03-23 R/processing-code/processingcode.R

@@ -25,9 +25,12 @@ library(here) #to set paths
 #package::function() that's not required one could just call the function
 #specifying the package makes it clearer where the function "lives",
 #but it adds typing. You can do it either way.
-data_location <- here::here("data","raw-data","500_Cities__Cancer__excluding_skin_cancer__among_adults_aged___18_years_20240314.csv")
+data_location <- here::here("data","raw-data","ObesityDataSet_raw_and_data_sinthetic.csv")
 rawdata <- read.csv(data_location)
 
+#check on codebook
+codebook <- here::here("data","raw-data","codebook.xlsx")
+read_excel(codebook)
 
 ## ---- exploredata --------
 #take a look at the data
@@ -51,23 +54,28 @@ skimr::skim(rawdata)
 #defined as a categorical variable but not a numeric variable.
 
 
-# choose needed categories
+# generated new variable
 d1 <- rawdata %>%
-  filter(GeographicLevel == "City", DataValueTypeID == "CrdPrv", StateAbbr != "US")
+  mutate(BMI = Weight / (Height)^2) %>%
+  rename(History = family_history_with_overweight, Water = CH2O, Alcohol = CALC, Obesity = NObeyesdad) %>%
+  mutate(Gender = if_else(Gender == "Male", 1, 2))
 
-# drop uninterested variables for a smaller dataset
-d2 <- d1 %>%
-  select(Year, StateAbbr, StateDesc, CityName, DataValueTypeID, Data_Value,PopulationCount)
+#rename variable and category
+d2 <- d1 %>% 
+  mutate(Alcohol = case_when(
+    Alcohol == "no" ~ 1,
+    Alcohol == "Sometimes" ~ 2,
+    Alcohol == "Frequently" ~ 3,
+    Alcohol == "Always" ~ 4,
+    TRUE ~ NA_integer_  # Handle any other cases
+  ))
+#Print the updated dataframe
+head(d2)
 
-# generating a new variable
+#Select variables
 d3 <- d2 %>%
-  mutate(Cancer_Case = round((Data_Value / 100) * PopulationCount))
-
-# re-check with the data
-dplyr::glimpse(d3)
-summary(d3)
+  select(Gender, Age, History, Water, Alcohol, FAF, MTRANS,BMI, Obesity)
 head(d3)
-skimr::skim(d3)
 
 ## ---- savedata --------
 # all done, data is clean now. 

### 2024-03-28 R/eda-code/eda.qmd
part 3 updated

@@
 ggsave(filename = figure_file, plot=bc1) 
+figure_file = here("results","figures","age-distribution.png")
+ggsave(filename = figure_file, plot=hp2) 
 figure_file = here("results","figures","obesity-distribution.png")
@@
 ggsave(filename = figure_file, plot=sbp4) 
+figure_file = here("results","figures","obesity level by age.png")
+ggsave(filename = figure_file, plot=sbp5)
 figure_file = here("results","figures","obesity level by family history.png")

### 2024-04-04 R/analysis-code/exploratory-analysis.qmd
part 4

@@
 lm1_result <- summary(lm1)
+lm1_result
 
@@
 
-Last, I will try to use all predictors that I believe may affect BMI to fit a model.
+I haven't include gender or age for model fitting, I will try to combine them in one model here.
 
 ```{r}
-mydata1 <- mydata %>%
-  select(History, Water, Alcohol, FAF, MTRANS, BMI)
-lm3 <- lm(BMI ~ ., data = mydata1)
-lm3_result <- summary(lm3)
+lm4 <- lm(BMI ~ Gender * Age, data = mydata)
+lm4_result <- summary(lm4)
+lm4_result
 
@@
 ```
+
+This linear model proves gender and age can also be predictors for BMI and there is interaction between them. So I will include them in the all predictor model.
+
+Last, I will try to use all predictors that I believe may affect BMI to fit a model.
+
+```{r}
+lm3 <- lm(BMI ~ Gender + Age + History + Water + Alcohol + FAF + MTRANS, data = mydata)
+lm3_result <- summary(lm3)
+lm3_result
+
+# save the result
+lm_file = here("results", "tables", "lm3table.rds")
+saveRDS(lm3_result, file = lm_file)
+```
+
+Choose predictors only important for the model fitting. Thanks to my classmates' replies, I received some useful suggestion for choosing predictors, which are chi-square, LASSO and lowest RMSE. 
+According to my searching result online, chi-square is good for categorical variable as outcome, and LASSO might shrink the coefficients too much.
+So I will go with the lowest RMSE in the following and see what will happen next.
+
+```{r}
+# Load necessary packages
+library(caret)
+
+# Split data into training and testing sets
+set.seed(123)  # Set seed for reproducibility
+trainIndex <- createDataPartition(mydata$BMI, p = 0.8, list = FALSE)
+trainData <- mydata[trainIndex, ]
+testData <- mydata[-trainIndex, ]
+
+# find important predictors without interactions
+
+# Define predictors
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+
+# Initialize variables to store results
+minRMSE1 <- Inf
+bestSubset1 <- NULL
+
+# Iterate through all possible combinations of predictors
+for (i in 1:length(predictors)) {
+  subsets <- combn(predictors, i)
+  for (j in 1:ncol(subsets)) {
+    predictors_subset <- subsets[, j]
+
+    # Fit model using subset of predictors
+    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
+    model <- lm(formula, data = trainData)
+
+    # Make predictions on test set
+    predictions <- predict(model, newdata = testData)
+
+    # Calculate RMSE
+    RMSE1 <- sqrt(mean((testData$BMI - predictions)^2))
+
+    # Update minRMSE and bestSubset if RMSE is lower
+    if (RMSE1 < minRMSE1) {
+      minRMSE1 <- RMSE1
+      bestSubset1 <- predictors_subset
+    }
+  }
+}
+
+# Print results
+print(paste("Best subset of predictors:", paste(bestSubset1, collapse = ", ")))
+print(paste("Lowest RMSE:", minRMSE1))
+```
+
+```{r}
+# Define predictors and response variable
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+interaction_terms <- c("Gender_Age", "FAF_Water")  # Define interaction terms
+response <- "BMI"
+
+# Create interaction terms in training and testing data
+trainData$Gender_Age <- trainData$Gender * trainData$Age
+trainData$FAF_Water <- trainData$FAF * trainData$Water
+
+testData$Gender_Age <- testData$Gender * testData$Age
+testData$FAF_Water <- testData$FAF * testData$Water
+
+# Update list of predictors to include interaction terms
+predictors_updated <- c(predictors, "Gender_Age", "FAF_Water")
+
+# Initialize variables to store results
+minRMSE2 <- Inf
+bestSubset2 <- NULL
+
+# Iterate through all possible combinations of predictors
+for (i in 1:length(predictors_updated)) {
+  subsets <- combn(predictors_updated, i)
+  for (j in 1:ncol(subsets)) {
+    predictors_subset <- subsets[, j]
+
+    # Fit model using subset of predictors
+    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
+    model <- lm(formula, data = trainData)
+
+    # Make predictions on test set
+    predictions <- predict(model, newdata = testData)
+
+    # Calculate RMSE
+    RMSE2 <- sqrt(mean((testData$BMI - predictions)^2))
+
+    # Update minRMSE and bestSubset if RMSE is lower
+    if (RMSE2 < minRMSE2) {
+      minRMSE2 <- RMSE2
+      bestSubset2 <- predictors_subset
+    }
+  }
+}
+
+# Print results
+print(paste("Best subset of predictors (including interactions):", paste(bestSubset2, collapse = ", ")))
+print(paste("Lowest RMSE (including interactions):", minRMSE2))
+```
+
+It is under my expectation that all-predictor model with the 2 interactions has the lowest RMSE. So I also would like to try LASSO regression model here and see the result.
+
+```{r}
+# Load necessary packages
+library(glmnet)
+library(caret)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Define predictors and response variable
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+response <- "BMI"
+
+# Prepare data
+x <- model.matrix(as.formula(paste(response, "~", paste(predictors, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
+y <- mydata[, response]
+
+# Fit LASSO regression model using cross-validation
+lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression
+
+# Extract optimal lambda value
+best_lambda <- lasso_model$lambda.min
+
+# Extract coefficients of predictors for the optimal lambda
+coef <- coef(lasso_model, s = best_lambda)
+
+# Identify non-zero coefficients (selected predictors)
+selected_predictors1 <- rownames(coef)[coef != 0]
+
+# Print selected predictors
+print(paste("Selected predictors:", paste(selected_predictors1, collapse = ", ")))
+```
+
+```{r}
+# Fit LASSO regression model using cross-validation with interaction terms
+x <- model.matrix(as.formula(paste(response, "~", paste(predictors_updated, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
+y <- mydata[, response]
+lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression
+
+# Extract optimal lambda value
+best_lambda <- lasso_model$lambda.min
+
+# Extract coefficients of predictors for the optimal lambda
+coef <- as.matrix(coef(lasso_model, s = "lambda.min"))
+
+# Identify non-zero coefficients (selected predictors)
+selected_predictors <- rownames(coef)[coef != 0]
+
+# Print selected predictors
+print(paste("Selected predictors (including interactions):", paste(selected_predictors, collapse = ", ")))
+```
+
+The LASSO regression result includes some specific levels in variables instead of the entire variables.
+
+Using CV to test evaluate the model with all predictors.
+
+```{r}
+# Define predictors and response variable
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+interaction_terms <- c("Gender_Age", "FAF_Water")
+response <- "BMI"
+
+# Define training control with repeated k-fold cross-validation
+train_control <- trainControl(method = "repeatedcv", 
+                              number = 10,       # Number of folds
+                              repeats = 3)       # Number of repetitions
+
+# Define the model formula
+formula <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))
+
+# Train the model using cross-validation
+model <- train(formula,
+               data = mydata,
+               method = "lm",                    # Use linear regression
+               trControl = train_control,       # Use defined training control
+               preProcess = c("center", "scale"),  # Preprocess data
+               metric = "RMSE")                    # Use RMSE as the metric
+
+# Print the cross-validation results
+print(model)
+```
+
 # Notes

### 2024-04-06 R/analysis-code/exploratory-analysis.qmd
part 4 update

@@
 First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.
+First I want to use number to represent the type of transportation. Walking requires most energy consumption, then it goes with public transportation and automobile is the least.
 
 ```{r}
-# Fit multinomial logistic regression model
-#library(nnet)
-#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)
+# Define a function to map transportation modes to numeric values
+map_transport <- function(transport_mode) {
+  if (transport_mode == "Walking") {
+    return(4)
+  } else if (transport_mode == "Motorbike") {
+    return(2)
+  } else if (transport_mode == "Public_Transportation") {
+    return(3)
+  } else if (transport_mode == "Automobile") {
+    return(1)
+  } else {
+    return(0)  # Return 0 for other cases
+  }
+}
 
-# Summary of the model
-#summary(multinom_model)
+# Apply the function to create the new variable
+mydata$MTRANSn <- sapply(mydata$MTRANS, map_transport)
+
+# Calculate correlation coefficient
+correlation <- cor(mydata$FAF, mydata$MTRANSn)
+
+# Print correlation coefficient
+print(correlation)
 ```
-I will try to figure it out later how I could make it fitting.
+Since the correlation coefficient is close to zero, there is almost no linear association between them. So this result doesn't really support the theory that I have heard from social media.
 
@@
 for (i in 1:length(predictors)) {
-  subsets <- combn(predictors, i)
-  for (j in 1:ncol(subsets)) {
-    predictors_subset <- subsets[, j]
+  subsets1 <- combn(predictors, i)
+  for (j in 1:ncol(subsets1)) {
+    predictors_subset1 <- subsets1[, j]
 
     # Fit model using subset of predictors
-    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
-    model <- lm(formula, data = trainData)
+    formula_lm1 <- paste("BMI ~", paste(predictors_subset1, collapse = " + "))
+    model_lm1 <- lm(formula_lm1, data = trainData)
 
     # Make predictions on test set
-    predictions <- predict(model, newdata = testData)
+    predictions_lm1 <- predict(model_lm1, newdata = testData)
 
     # Calculate RMSE
-    RMSE1 <- sqrt(mean((testData$BMI - predictions)^2))
+    RMSE1 <- sqrt(mean((testData$BMI - predictions_lm1)^2))
 
@@
       minRMSE1 <- RMSE1
-      bestSubset1 <- predictors_subset
+      bestSubset1 <- predictors_subset1
     }
@@
 for (i in 1:length(predictors_updated)) {
-  subsets <- combn(predictors_updated, i)
-  for (j in 1:ncol(subsets)) {
-    predictors_subset <- subsets[, j]
+  subsets2 <- combn(predictors_updated, i)
+  for (j in 1:ncol(subsets2)) {
+    predictors_subset2 <- subsets2[, j]
 
     # Fit model using subset of predictors
-    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
-    model <- lm(formula, data = trainData)
+    formula_lm2 <- paste("BMI ~", paste(predictors_subset2, collapse = " + "))
+    model_lm2 <- lm(formula_lm2, data = trainData)
 
     # Make predictions on test set
-    predictions <- predict(model, newdata = testData)
+    predictions_lm2 <- predict(model_lm2, newdata = testData)
 
     # Calculate RMSE
-    RMSE2 <- sqrt(mean((testData$BMI - predictions)^2))
+    RMSE2 <- sqrt(mean((testData$BMI - predictions_lm2)^2))
 
@@
       minRMSE2 <- RMSE2
-      bestSubset2 <- predictors_subset
+      bestSubset2 <- predictors_subset2
     }
@@
 
-# Fit LASSO regression model using cross-validation
-lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression
+# Define number of folds for cross-validation
+nfolds <- 10
 
-# Extract optimal lambda value
-best_lambda <- lasso_model$lambda.min
+# Perform cross-validation
+cv_model <- cv.glmnet(x, y, alpha = 1, nfolds = nfolds)  # Use alpha = 1 for LASSO regression
 
-# Extract coefficients of predictors for the optimal lambda
-coef <- coef(lasso_model, s = best_lambda)
-
-# Identify non-zero coefficients (selected predictors)
-selected_predictors1 <- rownames(coef)[coef != 0]
+# Extract selected predictors
+selected_predictors <- rownames(coef(cv_model, s = "lambda.min"))[-1]  # Exclude intercept column
 
 # Print selected predictors
-print(paste("Selected predictors:", paste(selected_predictors1, collapse = ", ")))
-```
-
-```{r}
-# Fit LASSO regression model using cross-validation with interaction terms
-x <- model.matrix(as.formula(paste(response, "~", paste(predictors_updated, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
-y <- mydata[, response]
-lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression
-
-# Extract optimal lambda value
-best_lambda <- lasso_model$lambda.min
-
-# Extract coefficients of predictors for the optimal lambda
-coef <- as.matrix(coef(lasso_model, s = "lambda.min"))
-
-# Identify non-zero coefficients (selected predictors)
-selected_predictors <- rownames(coef)[coef != 0]
-
-# Print selected predictors
-print(paste("Selected predictors (including interactions):", paste(selected_predictors, collapse = ", ")))
+print(paste("Selected predictors:", paste(selected_predictors, collapse = ", ")))
 ```
@@
 # Define training control with repeated k-fold cross-validation
-train_control <- trainControl(method = "repeatedcv", 
+train_control1 <- trainControl(method = "repeatedcv", 
                               number = 10,       # Number of folds
@@
 # Define the model formula
-formula <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))
+formula_CV <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))
 
 # Train the model using cross-validation
-model <- train(formula,
+model <- train(formula_CV,
                data = mydata,
@@
 
+I will also try to use obesity level which is categorical variable so I can use DA models to fit the data and might compare the result with linear models.
+There are totally 7 categories classified by the original authors of the dataset.
+
+```{r}
+library(MASS)
+library(caret)
+
+# Define the outcome variable
+outcome <- "Obesity"
+
+# Define predictor variables
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+
+# Create a training/testing split
+set.seed(123)  # Set seed for reproducibility
+trainIndex3 <- createDataPartition(mydata$Obesity, p = 0.8, list = FALSE)
+trainData3 <- mydata[trainIndex, ]
+testData3 <- mydata[-trainIndex, ]
+
+# Initialize variables to store results
+best_subset_DA <- NULL
+best_accuracy_DA <- 0
+
+# Iterate through all possible combinations of predictors
+for (i in 1:length(predictors)) {
+  subsets3 <- combn(predictors, i)
+  for (j in 1:ncol(subsets3)) {
+    predictors_subset_DA <- subsets3[, j]
+
+    # Fit the DA model using the subset of predictors
+    formula_DA <- as.formula(paste(outcome, "~", paste(predictors_subset_DA, collapse = " + ")))
+    model_DA <- lda(formula_DA, data = trainData3)
+
+    # Make predictions on the testing set
+    predicted_DA <- predict(model_DA, newdata = testData3)$class
+
+    # Calculate accuracy
+    accuracy_DA <- mean(predicted_DA == testData3$Obesity)
+
+    # Update best subset and accuracy if current accuracy is higher
+    if (accuracy_DA > best_accuracy_DA) {
+      best_subset_DA <- predictors_subset_DA
+      best_accuracy_DA <- accuracy_DA
+    }
+  }
+}
+
+# Print the best subset of predictors and accuracy
+print(paste("Best subset of predictors:", paste(best_subset_DA, collapse = ", ")))
+print(paste("Best accuracy:", best_accuracy_DA))
+```
+
+Adding 2 interactions to the DA model.
+
+```{r}
+# Define the outcome variable
+outcome <- "Obesity"
+
+# Update list of predictors to include interaction terms
+predictors_updated <- c(predictors, "Gender_Age", "FAF_Water")
+
+# Create a training/testing split
+set.seed(123)  # Set seed for reproducibility
+trainIndex4 <- createDataPartition(mydata$Obesity, p = 0.8, list = FALSE)
+trainData4 <- mydata[trainIndex, ]
+testData4 <- mydata[-trainIndex, ]
+
+# Create interaction terms in training and testing data
+trainData4$Gender_Age <- trainData4$Gender * trainData4$Age
+trainData4$FAF_Water <- trainData4$FAF * trainData4$Water
+
+testData4$Gender_Age <- testData4$Gender * testData4$Age
+testData4$FAF_Water <- testData4$FAF * testData4$Water
+
+# Initialize variables to store results
+best_subset_DA_U <- NULL
+best_accuracy_DA_U <- 0
+
+# Iterate through all possible combinations of predictors
+for (i in 1:length(predictors_updated)) {
+  subsets4 <- combn(predictors_updated, i)
+  for (j in 1:ncol(subsets4)) {
+    predictors_subset_DA_U <- subsets4[, j]
+
+    # Fit the DA model using the subset of predictors
+    formula_DA_U <- as.formula(paste(outcome, "~", paste(predictors_subset_DA_U, collapse = " + ")))
+    model_DA_U <- lda(formula_DA_U, data = trainData4)
+
+    # Make predictions on the testing set
+    predicted_DA_U <- predict(model_DA_U, newdata = testData4)$class
+
+    # Calculate accuracy
+    accuracy_DA_U <- mean(predicted_DA_U == testData4$Obesity)
+
+    # Update best subset and accuracy if current accuracy is higher
+    if (accuracy_DA_U > best_accuracy_DA_U) {
+      best_subset_DA_U <- predictors_subset_DA_U
+      best_accuracy_DA_U <- accuracy_DA_U
+    }
+  }
+}
+
+# Print the best subset of predictors and accuracy
+print(paste("Best subset of predictors:", paste(best_subset_DA_U, collapse = ", ")))
+print(paste("Best accuracy:", best_accuracy_DA_U))
+```
+After adding 2 interactions, the accuracy increased a little bit.
+
+Using cross validation to evaluate the model.
+
+```{r}
+# Define predictors and response variable
+mydata$Obesity <- factor(mydata$Obesity)
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+interaction_terms <- c("Gender_Age", "FAF_Water")
+outcome <- "Obesity"
+
+# Define the model formula
+formula_CV_DA <- as.formula(paste(outcome, "~", paste(c(predictors, interaction_terms), collapse = " + ")))
+
+# Define training control with repeated k-fold cross-validation
+train_control2 <- trainControl(method = "repeatedcv", 
+                               number = 10,            # Number of folds
+                               repeats = 3,            # Number of repetitions
+                               classProbs = TRUE,      # Compute class probabilities
+                               summaryFunction = twoClassSummary)  # Use two-class summary
+
+# Fit the DA model using cross-validation
+lda_model <- train(formula = formula_CV_DA,
+                   data = mydata,
+                   method = "lda",             # Use LDA method
+                   trControl = train_control2,  # Use defined training control
+                   metric = "Accuracy",        # Use accuracy as the metric
+                   preProcess = c("center", "scale")) # Preprocess data
+
+print(lda_model)
+```
+
 # Notes

### 2024-04-07 R/analysis-code/exploratory-analysis.qmd
part 4 manuscript update

@@
 library(ggplot2)
+library(tidymodels)
+library(discrim)
 ```
@@
 print(paste("Lowest RMSE:", minRMSE1))
+
+# calculate null model
+
+# Compute the mean BMI from the training data
+mean_bmi_train <- mean(trainData$BMI)
+
+# Create a vector of predicted values equal to the mean BMI from training data
+predicted_bmi_test <- rep(mean_bmi_train, nrow(testData))
+
+# Compute RMSE (Root Mean Squared Error) on test data
+rmse_null_test <- sqrt(mean((testData$BMI - predicted_bmi_test)^2))
+
+# Print the results
+print(paste("Null Model RMSE on Test Data:", rmse_null_test))
+```
+```{r}
+# Create a data frame for plotting
+plot_data1 <- data.frame(Observed = testData$BMI, 
+                        Predicted_All = predictions_lm1,
+                        Predicted_Null = predicted_bmi_test)
+
+# Plot using ggplot
+p1 <- ggplot(plot_data1, aes(x = Observed, y = Predicted_All)) +
+  geom_point(color = "blue", size = 2, shape = 16) +
+  geom_point(aes(y = Predicted_Null), color = "green", size = 2, shape = 16) +
+  geom_abline(intercept = 0, slope = 1, color = "red") +
+  labs(x = "Observed BMI", y = "Predicted BMI", title = "Observed vs Predicted BMI") +
+  theme_minimal() +
+  theme(legend.position = "top") +
+  scale_shape_manual(values = c(16, 16)) +
+  scale_color_manual(values = c("blue", "green")) +
+  guides(color = guide_legend(title = "Model"))
+plot(p1)
+
+# save the plot
+figure_file = here("results","figures","Observed-Predicted1 .png")
+ggsave(filename = figure_file, plot=p1) 
 ```
@@
 
+```{r}
+# Create a data frame for plotting
+plot_data2 <- data.frame(Observed = testData$BMI, 
+                        Predicted_All = predictions_lm2,
+                        Predicted_Null = predicted_bmi_test)
+
+# Plot using ggplot
+p2 <- ggplot(plot_data2, aes(x = Observed, y = Predicted_All)) +
+  geom_point(color = "red", size = 2, shape = 16) +
+  geom_point(aes(y = Predicted_Null), color = "yellow", size = 2, shape = 16) +
+  geom_abline(intercept = 0, slope = 1, color = "black") +
+  labs(x = "Observed BMI", y = "Predicted BMI", title = "Observed vs Predicted BMI with interactions") +
+  theme_minimal() +
+  theme(legend.position = "top") +
+  scale_shape_manual(values = c(16, 16)) +
+  scale_color_manual(values = c("red", "yellow")) +
+  guides(color = guide_legend(title = "Model"))
+plot(p2)
+# save the plot
+figure_file = here("results","figures","Observed-Predicted2 .png")
+ggsave(filename = figure_file, plot=p2) 
+```
+
 It is under my expectation that all-predictor model with the 2 interactions has the lowest RMSE. So I also would like to try LASSO regression model here and see the result.
@@
 
+# Create interaction terms in data
+mydata$Gender_Age <- mydata$Gender * mydata$Age
+mydata$FAF_Water <- mydata$FAF * mydata$Water
+
 # Define training control with repeated k-fold cross-validation
@@
                method = "lm",                    # Use linear regression
-               trControl = train_control,       # Use defined training control
+               trControl = train_control1,       # Use defined training control
                preProcess = c("center", "scale"),  # Preprocess data
@@
 
-# Define training control with repeated k-fold cross-validation
-train_control2 <- trainControl(method = "repeatedcv", 
-                               number = 10,            # Number of folds
-                               repeats = 3,            # Number of repetitions
-                               classProbs = TRUE,      # Compute class probabilities
-                               summaryFunction = twoClassSummary)  # Use two-class summary
+# Set seed for reproducibility
+set.seed(1234)
 
-# Fit the DA model using cross-validation
-lda_model <- train(formula = formula_CV_DA,
-                   data = mydata,
-                   method = "lda",             # Use LDA method
-                   trControl = train_control2,  # Use defined training control
-                   metric = "Accuracy",        # Use accuracy as the metric
-                   preProcess = c("center", "scale")) # Preprocess data
+# Create 10-fold cross-validation folds
+new_folds <- vfold_cv(mydata, v = 10)
 
-print(lda_model)
+# Define model specification
+lda_spec <- discrim_linear() %>%
+  set_engine("MASS") %>%
+  set_mode("classification")
+
+# Workflow for model with specified predictors
+lda_model <- workflow() %>%
+  add_model(lda_spec) %>%
+  add_formula(formula_CV_DA)
+
+# Workflow for model using all predictors
+lda_wf <- workflow() %>%
+  add_model(lda_spec) %>%
+  add_formula(formula_CV_DA)
+
+# Perform CV model training and evaluation
+lda_CV <- lda_wf %>%
+  fit_resamples(resamples = new_folds, metrics = metric_set(accuracy))
+
+# Collect RMSE values
+collect_metrics(lda_CV)
 ```

### 2024-04-07 products/manuscript/Manuscript copy.qmd
part 3 updated
part 4
part 4 manuscript update

@@
 
-In this project, I have found......
+Obesity is a chronic complex disease which may increase the risk of severe health issues and even cancers.
+And the condition of obesity in real life could be related to physical behaviors.
+The project is exploring the potential factors that might be related to obesity level using machine learning models.
+The data analysis result has shown that age, gender, family history, drinking water, physical activity, and commute transportation are all predictors for obesity level, which is represented by individual BMI.
+As for the outcome, when linear model is fit to data which has BMI as continuous variable, the performance is better than using obesity level as categorical variable.
 
@@
 
+I made a code book based on the article reporting this dataset [@PALECHOR2019104344] first before doing any cleaning.
+
 Obesity level is a categorical variable defined by BMI range, so BMI as a new variable is created and it could be analyzed as a numeric variable.
@@
 Figure 2.
+Age distribution
+
+![](images/age-distribution.png)
+
+Figure 3.
 Obesity level distribution.
@@
 
-Figure 3.
+Figure 4.
 Gender distribution in each obesity levels.
@@
 
-Figure 4.
+Figure 5.
+Age distribution in each obesity levels.
+
+![](images/obesity%20level%20by%20age.png)
+
+Figure 6.
 Violin plot of BMI by family obesity history.
 
-![](images/obesity level by family history-02.png)
+![](images/obesity%20level%20by%20family%20history-01.png)
 
@@
 
-Since some obvious relationship could be shown in figures, I choose other variables as predictors in exploratory analysis
+Since some obvious relationship could be shown in figures, I choose other variables as predictors in exploratory analysis.
 
-First, I would like to fit linear model for BMI by both water consumption and physical activity.
+First, I was curious that whether there is a negative correlation between transportation type and physical activity frequency, I converted transportation types into numbers indicating the energy consuming levels from 1 (lowest automobile) to 4 (highest walking).
+Then I used correlation function to calculate the coefficient for these two variable but it was pretty low (0.0059).
+
+Then, I would like to fit linear model for BMI by both water consumption and physical activity.
 And the result indicates that there is a connection effect between them two, which means that both of them contribution to obesity control.
@@
 
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
+Since I found there were interactions between physical activity-drinking water and age-gender, then I chose BMI as outcome and fitted the data with linear regression model.
+I asked chatGPT for help to write code and used the code to find the best combination of all the predictors included in the processed dataset.
+By comparing RMSE values of each combination, all predictors were selected.
+After that, I calculated RMSE of the null model (8.04) and it was a little bit higher than all predictor model (6.53).
+
+![](images/Observed-Predicted1 .png){width="385"}
+
+According to the suggestions I received from module 11's discussion, I also included 2 interactions as predictors and used linear model to fit data again, it was under my expectation that it lowered the RMSE (6.39), but didn't make much difference in fact.
+
+![](images/Observed-Predicted2 .png){width="385"}
+
+Then I used cross validation to check if the all predictors model overfits the data and the average RMSE I got was 6.32, which was not so different from data splitting.
+
+And I also tried to fit LASSO model with all predictors inspired by the reply in discord, but I couldn't interpret the result well because it selected specific categories rather than the whole variables as the predictors as the best combination, which I couldn't figure out the meaning of that.
+So I only considered the linear regression result.
+
+Because I had a question for whether using BMI as an outcome is preferable than obesity level, I used discriminant analysis model to fit the data with obesity level as outcome inspired by module 12 discussion since there were 7 categories in total under obesity level.
+I got the accuracy for all predictors without interactions was 0.45, and 0.48 with interactions.
+The average accuracy of cross validation result was 0.48, almost as same as the result with interactions included.
 
@@
 
-*Summarize what you did, what you found and what it means.*
+I chose a dataset that summarize an investigation result done in South American countries about individual physical behaviors and the classification of obesity level based on the information of height and weight obtained from the investigation.
+I had some questions to explore and did data analysis on this dataset.
+
+I found simple linear model fitting can answer my questions, which are that alcohol consumption is correlated to obesity, as well as there was interactions between physical activity frequency and drinking water.
+Besides, I found there was nearly no correlation between transportation choice and physical activity, which could answer another question I had.
+
+For machine learning model fitting, I found the best predictors for BMI index based on the value of some performance metrics after trying both linear regression and classification model.
+It is hard to compare the metrics between two models, but I tend to choose linear regression model and BMI as outcome since RMSE at value 6 means relatively low and the model fits well for me compared to accuracy at 0.48 means the model can predict 48% of the test data correctly.
+And the result of cross validation shows a a sign of not overfitting.
+
+So I think I would choose linear regression as the better prediction model and all variables
 
@@
 
-*Discuss what you perceive as strengths and limitations of your analysis.*
+I think my analysis has include two different types of model according to the type of outcome, which can provide more information and the result can be shown in different aspects.
+
+But there is still a concern that whether my model is overfitting not not since the RMSE I have got is much lower than exercise 8 which is several hundreds even if I have applied cross validation.
+But I still cannot find a good way to figure it out based on my knowledge so far.
 
@@
 
-**Haven't updated yet but will do it later!!!**
-
-*What are the main take-home messages?*
-
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
+The take-home message for this project is that obesity is correlated to individual physical behaviors.
+Based on the result of data analysis, age, gender, family history, drinking water, physical activity, and commute transportation are all predictors for obesity level which is represented by BMI index.
 

### 2024-04-22 R/analysis-code/exploratory-analysis.qmd
part 5

--- 2024-04-04 R/analysis-code/exploratory-analysis.qmd
@@
 library(ggplot2)
+library(tidymodels)
+library(discrim)
+library(randomForest)
+library(caret)
+library(lattice)
+library(tidymodels)
+library(ranger)
+library(parsnip)
+library(glmnet)
+library(MASS)
+library(nnet)
 ```
@@
 First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.
+First I want to use number to represent the type of transportation.
+Walking requires most energy consumption, then it goes with public transportation and automobile is the least.
 
 ```{r}
-# Fit multinomial logistic regression model
-#library(nnet)
-#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)
+# Define a function to map transportation modes to numeric values
+map_transport <- function(transport_mode) {
+  if (transport_mode == "Walking") {
+    return(4)
+  } else if (transport_mode == "Motorbike") {
+    return(2)
+  } else if (transport_mode == "Public_Transportation") {
+    return(3)
+  } else if (transport_mode == "Automobile") {
+    return(1)
+  } else {
+    return(0)  # Return 0 for other cases
+  }
+}
 
-# Summary of the model
-#summary(multinom_model)
+# Apply the function to create the new variable
+mydata$MTRANSn <- sapply(mydata$MTRANS, map_transport)
+
+# Calculate correlation coefficient
+correlation <- cor(mydata$FAF, mydata$MTRANSn)
+
+# Print correlation coefficient
+print(correlation)
 ```
-I will try to figure it out later how I could make it fitting.
+
+Since the correlation coefficient is close to zero, there is almost no linear association between them. So this result doesn't really support the theory that I have heard from social media.
 
@@
 lm1_result <- summary(lm1)
-lm1_result
 
@@
 ```
+
 I think it makes a little bit sense to me that there is a connection effect of water and physical activity on BMI, which might be supportive for what we learn that drinking more water and doing more exercise will lower the risk of obesity.
@@
 lm4_result <- summary(lm4)
-lm4_result
 
 # save the result
-lm_file = here("results", "tables", "lm3table.rds")
-saveRDS(lm3_result, file = lm_file)
+lm_file = here("results", "tables", "lm4table.rds")
+saveRDS(lm4_result, file = lm_file)
 ```
@@
 lm3_result <- summary(lm3)
-lm3_result
 
@@
 
-Choose predictors only important for the model fitting. Thanks to my classmates' replies, I received some useful suggestion for choosing predictors, which are chi-square, LASSO and lowest RMSE. 
+Choose predictors only important for the model fitting.
+Thanks to my classmates' replies, I received some useful suggestion for choosing predictors, which are chi-square, LASSO and lowest RMSE.
 According to my searching result online, chi-square is good for categorical variable as outcome, and LASSO might shrink the coefficients too much.
@@
 ```{r}
-# Load necessary packages
-library(caret)
-
 # Split data into training and testing sets
@@
 testData <- mydata[-trainIndex, ]
+# Create indices for 10-fold cross-validation
+cv_folds <- createFolds(trainData$BMI, k = 10)
+```
 
-# find important predictors without interactions
-
+```{r}
+# predictors excluding interactions first
 # Define predictors
@@
 
-# Initialize variables to store results
-minRMSE1 <- Inf
-bestSubset1 <- NULL
-
 # Iterate through all possible combinations of predictors
 for (i in 1:length(predictors)) {
-  subsets <- combn(predictors, i)
-  for (j in 1:ncol(subsets)) {
-    predictors_subset <- subsets[, j]
+  subsets1 <- combn(predictors, i)
+  for (j in 1:ncol(subsets1)) {
+    predictors_subset1 <- subsets1[, j]
+
+    # Initialize variable to store RMSE for current combination of predictors
+    total_rmse_lm1 <- 0
+
+    # Perform cross-validation
+    for (fold in 1:10) {
+      # Extract indices for training and validation data for current fold
+      train_indices <- unlist(cv_folds[-fold])
+      validation_indices <- cv_folds[[fold]]
 
       # Fit model using subset of predictors
-    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
-    model <- lm(formula, data = trainData)
+      formula_lm1 <- paste("BMI ~", paste(predictors_subset1, collapse = " + "))
+      model_lm1 <- lm(formula_lm1, data = trainData[train_indices, ])
 
-    # Make predictions on test set
-    predictions <- predict(model, newdata = testData)
+      # Make predictions on validation set
+      predictions_lm1 <- predict(model_lm1, newdata = trainData[validation_indices, ])
 
-    # Calculate RMSE
-    RMSE1 <- sqrt(mean((testData$BMI - predictions)^2))
+      # Calculate RMSE for current fold
+      fold_rmse_lm1 <- sqrt(mean((trainData$BMI[validation_indices] - predictions_lm1)^2))
 
-    # Update minRMSE and bestSubset if RMSE is lower
-    if (RMSE1 < minRMSE1) {
-      minRMSE1 <- RMSE1
-      bestSubset1 <- predictors_subset
+      # Add RMSE for current fold to total RMSE
+      total_rmse_lm1 <- total_rmse_lm1 + fold_rmse_lm1
+    }
+
+    # Calculate average RMSE across all folds
+    avg_rmse_lm1 <- total_rmse_lm1 / 10
+
+    # Update minRMSE and bestSubset if average RMSE is lower
+    if (avg_rmse_lm1 < minRMSE1) {
+      minRMSE1 <- avg_rmse_lm1
+      bestSubset1 <- predictors_subset1
     }
@@
 ```{r}
+# calculate null model RMSE
+# Initialize variable to store total RMSE across all folds
+total_rmse_null <- 0
+
+# Perform cross-validation
+for (fold in 1:10) {
+  # Extract indices for training and validation data for current fold
+  train_indices <- unlist(cv_folds[-fold])
+  validation_indices <- cv_folds[[fold]]
+
+  # Compute the mean BMI from the training data
+  mean_bmi_train <- mean(trainData$BMI[train_indices])
+
+  # Create a vector of predicted values equal to the mean BMI from training data
+  predicted_bmi_validation <- rep(mean_bmi_train, length(validation_indices))
+
+  # Compute RMSE (Root Mean Squared Error) on validation data for current fold
+  rmse_null_validation <- sqrt(mean((trainData$BMI[validation_indices] - predicted_bmi_validation)^2))
+
+  # Add RMSE for current fold to total RMSE
+  total_rmse_null <- total_rmse_null + rmse_null_validation
+}
+
+# Calculate average RMSE across all folds
+avg_rmse_null <- total_rmse_null / 10
+
+# Print the results
+print(paste("Null Model RMSE using Cross-Validation:", avg_rmse_null))
+```
+
+```{r}
+# interactions included in predictors
 # Define predictors and response variable
@@
 for (i in 1:length(predictors_updated)) {
-  subsets <- combn(predictors_updated, i)
-  for (j in 1:ncol(subsets)) {
-    predictors_subset <- subsets[, j]
+  subsets2 <- combn(predictors_updated, i)
+  for (j in 1:ncol(subsets2)) {
+    predictors_subset2 <- subsets2[, j]
+
+    # Initialize variable to store RMSE for current combination of predictors
+    total_rmse_lm2 <- 0
+
+    # Perform cross-validation
+    for (fold in 1:10) {
+      # Extract indices for training and validation data for current fold
+      train_indices <- unlist(cv_folds[-fold])
+      validation_indices <- cv_folds[[fold]]
 
       # Fit model using subset of predictors
-    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
-    model <- lm(formula, data = trainData)
+      formula_lm2 <- paste("BMI ~", paste(predictors_subset2, collapse = " + "))
+      model_lm2 <- lm(formula_lm2, data = trainData[train_indices, ])
 
-    # Make predictions on test set
-    predictions <- predict(model, newdata = testData)
+      # Make predictions on validation set
+      predictions_lm2 <- predict(model_lm2, newdata = trainData[validation_indices, ])
 
-    # Calculate RMSE
-    RMSE2 <- sqrt(mean((testData$BMI - predictions)^2))
+      # Calculate RMSE for current fold
+      fold_rmse_lm2 <- sqrt(mean((trainData$BMI[validation_indices] - predictions_lm2)^2))
 
-    # Update minRMSE and bestSubset if RMSE is lower
-    if (RMSE2 < minRMSE2) {
-      minRMSE2 <- RMSE2
-      bestSubset2 <- predictors_subset
+      # Add RMSE for current fold to total RMSE
+      total_rmse_lm2 <- total_rmse_lm2 + fold_rmse_lm2
+    }
+
+    # Calculate average RMSE across all folds
+    avg_rmse_lm2 <- total_rmse_lm2 / 10
+
+    # Update minRMSE and bestSubset if average RMSE is lower
+    if (avg_rmse_lm2 < minRMSE2) {
+      minRMSE2 <- avg_rmse_lm2
+      bestSubset2 <- predictors_subset2
     }
@@
 # Print results
-print(paste("Best subset of predictors (including interactions):", paste(bestSubset2, collapse = ", ")))
-print(paste("Lowest RMSE (including interactions):", minRMSE2))
+print(paste("Best subset of predictors (including interactions) using cross-validation:", paste(bestSubset2, collapse = ", ")))
+print(paste("Lowest RMSE (including interactions) using cross-validation:", minRMSE2))
 ```
 
-It is under my expectation that all-predictor model with the 2 interactions has the lowest RMSE. So I also would like to try LASSO regression model here and see the result.
+It shows different predictor combinations but the good thing is both of the linear models have lower RMSE than null model which indicates they perform better than null model.
+Then, I will try LASSO model.
 
 ```{r}
-# Load necessary packages
-library(glmnet)
-library(caret)
+# Initialize variables to store results
+min_rmse_lasso <- Inf
+best_subset_lasso <- NULL
 
-# Set seed for reproducibility
-set.seed(123)
+# Perform LASSO regression with cross-validation
+lasso_model <- cv.glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, nfolds = 10)
 
-# Define predictors and response variable
-predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
-response <- "BMI"
+# Extract the optimal lambda value
+optimal_lambda <- lasso_model$lambda.min
 
-# Prepare data
-x <- model.matrix(as.formula(paste(response, "~", paste(predictors, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
-y <- mydata[, response]
+# Fit the LASSO model using the optimal lambda
+lasso_model_optimal <- glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, lambda = optimal_lambda)
 
-# Fit LASSO regression model using cross-validation
-lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression
+# Compute RMSE on test data using the optimal LASSO model
+predictions_lasso <- predict(lasso_model_optimal, newx = as.matrix(testData[, predictors_updated]))
+rmse_lasso <- sqrt(mean((testData$BMI - predictions_lasso)^2))
 
-# Extract optimal lambda value
-best_lambda <- lasso_model$lambda.min
+# Extract coefficients of the LASSO model
+lasso_coefficients <- as.matrix(coef(lasso_model_optimal))
 
-# Extract coefficients of predictors for the optimal lambda
-coef <- coef(lasso_model, s = best_lambda)
+# Print results
+cat("Optimal lambda:", optimal_lambda, "\n")
+cat("RMSE with LASSO:", rmse_lasso, "\n")
 
-# Identify non-zero coefficients (selected predictors)
-selected_predictors1 <- rownames(coef)[coef != 0]
+# Print non-zero coefficients (selected predictors)
+selected_predictors <- rownames(lasso_coefficients)[lasso_coefficients != 0, drop = FALSE]
+print("Selected predictors by LASSO:")
+print(selected_predictors)
+```
 
-# Print selected predictors
-print(paste("Selected predictors:", paste(selected_predictors1, collapse = ", ")))
+After that, I will select another regression model to fit data, here I will use random forest model.
+
+```{r}
+# Create formula with response and all predictors
+predictors_updated <- c(predictors, "Gender_Age", "FAF_Water")
+formula_rf <- as.formula(paste(response, "~", paste(predictors_updated, collapse = " + ")))
+
+# Initialize variables to store results
+best_rmse_rf <- Inf
+best_predictors_rf <- NULL
+
+# Forward selection with cross-validation
+for (fold in seq_along(cv_folds)) {
+  # Extract training and validation indices for current fold
+  train_indices <- unlist(cv_folds[-fold])
+  validation_indices <- cv_folds[[fold]]
+
+  # Initialize variables to store results for current fold
+  best_rmse_fold <- Inf
+  best_predictors_fold <- NULL
+
+  # Forward selection
+  for (i in seq_along(all_predictors)) {
+    predictors_subset_rf <- all_predictors[1:i]
+    formula_subset_rf <- as.formula(paste(response, "~", paste(predictors_subset_rf, collapse = " + ")))
+
+    # Fit random forest model using the training data for the current fold
+    model_rf <- randomForest(formula = formula_subset_rf, data = mydata[train_indices, ])
+
+    # Make predictions on the validation data for the current fold
+    predictions_rf <- predict(model_rf, newdata = mydata[validation_indices, ])
+
+    # Calculate RMSE for the current fold
+    rmse_rf <- sqrt(mean((mydata$BMI[validation_indices] - predictions_rf)^2))
+
+    # Check if current RMSE is better than the best RMSE so far for this fold
+    if (rmse_rf < best_rmse_fold) {
+      best_rmse_fold <- rmse_rf
+      best_predictors_fold <- predictors_subset_rf
+    }
+  }
+
+  # Check if the best RMSE for this fold is better than the best RMSE so far
+  if (best_rmse_fold < best_rmse_rf) {
+    best_rmse_rf <- best_rmse_fold
+    best_predictors_rf <- best_predictors_fold
+  }
+}
+
+# Print the best combination of predictors
+print("Best combination of predictors for random forest model:")
+print(best_predictors_rf)
+print(best_rmse_rf)
+```
+
+Since the measurement of obesity also can be shown by different degree levels, I will use obesity level as a new outcome and fit logistic models to the dataset so that more possibilities can be listed and more models can be compared.
+
+First, I can use discriminant analysis model to fit the data since here are totally 7 categories classified by the original authors of the dataset.
+
+```{r}
+# Define the outcome variable
+outcome <- "Obesity"
+
+# Define predictor variables
+predictors_updated <- c(predictors, "Gender_Age", "FAF_Water")
+
+# Initialize variables to store results
+best_subset_DA <- NULL
+best_accuracy_DA <- 0
+
+# Create cross-validation folds
+cv_fold1 <- createFolds(trainData$Obesity, k = 10)
+
+# Iterate through all possible combinations of predictors
+for (i in 1:length(predictors_updated)) {
+  subsets <- combn(predictors_updated, i)
+  for (j in 1:ncol(subsets)) {
+    predictors_subset_DA <- subsets[, j]
+
+    # Initialize variable to store accuracy for each fold
+    fold_accuracies <- numeric(length(cv_fold1))
+
+    # Perform cross-validation
+    for (fold_index in seq_along(cv_fold1)) {
+      # Extract training and validation indices for current fold
+      train_indices <- unlist(cv_fold1[-fold_index])
+      validation_indices <- cv_fold1[[fold_index]]
+
+      # Fit the DA model using the subset of predictors on the training data for the current fold
+      formula_DA <- as.formula(paste(outcome, "~", paste(predictors_subset_DA, collapse = " + ")))
+      model_DA <- lda(formula_DA, data = mydata[train_indices, ])
+
+      # Make predictions on the validation set
+      predicted_DA <- predict(model_DA, newdata = trainData[validation_indices, ])
+
+      # Calculate accuracy for the current fold
+      fold_accuracies[fold_index] <- mean(predicted_DA$class == mydata$Obesity[validation_indices])
+    }
+
+    # Calculate average accuracy across all folds
+    average_accuracy <- mean(fold_accuracies)
+
+    # Update best subset and accuracy if current accuracy is higher
+    if (average_accuracy > best_accuracy_DA) {
+      best_subset_DA <- predictors_subset_DA
+      best_accuracy_DA <- average_accuracy
+    }
+  }
+}
+
+# Print the best subset of predictors and accuracy
+print(paste("Best subset of predictors:", paste(best_subset_DA, collapse = ", ")))
+print(paste("Best average accuracy:", best_accuracy_DA))
+```
+
+I also need to know the null model result for this dataset.
+
+```{r}
+# Calculate the frequency of each category in the outcome variable
+outcome_frequency <- table(trainData$Obesity)
+
+# Find the most frequent category
+most_frequent_category <- names(outcome_frequency)[which.max(outcome_frequency)]
+
+# Create a vector of predicted values with the most frequent category
+predicted_outcome <- rep(most_frequent_category, nrow(trainData))
+
+# Calculate the accuracy of the null model
+accuracy_null_model <- sum(predicted_outcome == trainData$Obesity) / length(trainData$Obesity)
+
+# Print the accuracy
+print(paste("Accuracy of the null model:", accuracy_null_model))
+```
+
+Then, I choose another model that can also handle multicategorical outcome named Multinominal Logistic Regression.
+
+```{r}
+# Define outcome variable
+outcome <- "Obesity"
+
+# Create formula
+formula_mlr <- as.formula(paste(outcome, "~", paste(predictors_updated, collapse = " + ")))
+
+# Initialize variables to store results
+best_accuracy_mlr <- 0
+best_predictors_mlr <- NULL
+
+# Perform cross-validation
+for (fold in 1:10) {
+  # Extract indices for training and validation data for current fold
+  train_indices <- unlist(cv_fold1[-fold])
+  validation_indices <- cv_fold1[[fold]]
+
+  # Fit the Multinomial Logistic Regression model
+  model_mlr <- multinom(formula_mlr, data = trainData[train_indices, ])
+
+  # Make predictions on the validation set
+  predicted_mlr <- predict(model_mlr, newdata = trainData[validation_indices, ], type = "class")
+
+  # Calculate accuracy
+  accuracy_mlr <- mean(predicted_mlr == trainData$Obesity[validation_indices])
+
+  # Update best_accuracy and best_predictors if current accuracy is higher
+  if (accuracy_mlr > best_accuracy_mlr) {
+    best_accuracy_mlr <- accuracy_mlr
+    best_predictors_mlr <- predictors_updated
+  }
+}
+
+# Print the best predictor combination and highest accuracy
+print(paste("Best predictor combination:", paste(best_predictors_mlr, collapse = ", ")))
+print(paste("Highest Accuracy:", best_accuracy_mlr))
+```
+
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Model                             | Predictor                                 | Metric Value  | Metric Value on testing data |
++===================================+===========================================+===============+==============================+
+| Linear null                       | \-                                        | RMSE=8.00     | RMSE=8.04                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Linear regression 1               | Age, History, Alcohol, MTRANS             | RMSE=0.62     | RMSE=6.61                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Linear regression 2               | All                                       | RMSE=6.30     | RMSE=3.99                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Linear regression 2 + LASSO model | All                                       | RMSE=7.38     | RMSE=7.38                    |
+|                                   |                                           |               |                              |
+|                                   |                                           | l=0.0011      |                              |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Random forest                     | Gender, Age, History, Water, Alcohol, FAF | RMSE=3.22     | RMSE=3.97                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Logistic null                     | \-                                        | Accuracy=0.17 |                              |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Discriminant analysis             | All                                       | Accuracy=0.50 | Accuracy=0.48                |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Multinominal logistic regression  | All                                       | Accuracy=0.58 | Accuracy=0.52                |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+
+Then I will use testData to assess the quality to the model that I choose.
+
+```{r}
+# Define formula for random forest
+formula_rf_test <- as.formula(paste("BMI ~", paste(predictors_updated, collapse = " + ")))
+
+# Fit random forest model using the entire training data
+model_rf_test <- randomForest(formula = formula_rf_test, data = trainData)
+
+# Make predictions on the test data
+predictions_rf_test <- predict(model_rf_test, newdata = testData)
+
+# Calculate RMSE on test data
+rmse_rf_test <- sqrt(mean((testData$BMI - predictions_rf_test)^2))
+
+# Print RMSE on test data
+print(rmse_rf_test)
+
+# Null model of test data
+# Compute the mean BMI from the training data
+mean_bmi_train <- mean(trainData$BMI)
+
+# Create a vector of predicted values equal to the mean BMI from training data for the test data
+predicted_bmi_test <- rep(mean_bmi_train, nrow(testData))
+
+# Calculate RMSE (Root Mean Squared Error) on test data
+rmse_null_test <- sqrt(mean((testData$BMI - predicted_bmi_test)^2))
+
+# Print the results
+print(paste("Null Model RMSE on Test Data:", rmse_null_test))
+
+# Create a data frame for plotting
+plot_data_rf <- data.frame(Observed = testData$BMI, 
+                           Predicted_RF = predictions_rf_test,
+                           Predicted_Null = predicted_bmi_test)
+
+# Plot using ggplot
+p_rf <- ggplot(plot_data_rf, aes(x = Observed, y = Predicted_RF)) +
+  geom_point(color = "blue", size = 2, shape = 16) +
+  geom_point(aes(y = Predicted_Null), color = "green", size = 2, shape = 16) +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(x = "Observed BMI", y = "Predicted BMI", title = "Observed vs Predicted BMI") +
+  theme_minimal() +
+  theme(legend.position = "top") +
+  scale_shape_manual(values = c(16, 16)) +
+  scale_color_manual(values = c("blue", "green")) +
+  guides(color = guide_legend(title = "Model"))
+
+plot(p_rf)
+
+# save the plot
+figure_file = here("results","figures","Observed-Predicted .png")
+ggsave(filename = figure_file, plot=p_rf) 
 ```
@@
 ```{r}
-# Fit LASSO regression model using cross-validation with interaction terms
-x <- model.matrix(as.formula(paste(response, "~", paste(predictors_updated, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
-y <- mydata[, response]
-lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression
+# Define formula for mlr model
+formula_mlr_test <- as.formula(paste("Obesity ~", paste(predictors_updated, collapse = " + ")))
 
-# Extract optimal lambda value
-best_lambda <- lasso_model$lambda.min
+# Fit the Multinomial Logistic Regression model using the entire training data
+model_mlr_test <- multinom(formula_mlr_test, data = trainData)
 
-# Extract coefficients of predictors for the optimal lambda
-coef <- as.matrix(coef(lasso_model, s = "lambda.min"))
+# Make predictions on the test data
+predicted_mlr_test <- predict(model_mlr_test, newdata = testData, type = "class")
 
-# Identify non-zero coefficients (selected predictors)
-selected_predictors <- rownames(coef)[coef != 0]
+# Calculate accuracy on the test data
+accuracy_mlr_test <- mean(predicted_mlr_test == testData$Obesity)
 
-# Print selected predictors
-print(paste("Selected predictors (including interactions):", paste(selected_predictors, collapse = ", ")))
+# Print the accuracy
+print(paste("Accuracy of the MLR model on test data:", accuracy_mlr_test))
 ```
 
-The LASSO regression result includes some specific levels in variables instead of the entire variables.
-
-Using CV to test evaluate the model with all predictors.
+Here I try to assess every model I choose and check the result.
 
 ```{r}
-# Define predictors and response variable
-predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
-interaction_terms <- c("Gender_Age", "FAF_Water")
-response <- "BMI"
+# Compute the mean BMI from the training data
+mean_bmi_training <- mean(trainData$BMI)
 
-# Define training control with repeated k-fold cross-validation
-train_control <- trainControl(method = "repeatedcv", 
-                              number = 10,       # Number of folds
-                              repeats = 3)       # Number of repetitions
+# Create a vector of predicted values equal to the mean BMI from training data
+predicted_bmi_test <- rep(mean_bmi_train, nrow(testData))
 
-# Define the model formula
-formula <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))
+# Calculate RMSE on test data
+rmse_null_test <- sqrt(mean((testData$BMI - predicted_bmi_test)^2))
 
-# Train the model using cross-validation
-model <- train(formula,
-               data = mydata,
-               method = "lm",                    # Use linear regression
-               trControl = train_control,       # Use defined training control
-               preProcess = c("center", "scale"),  # Preprocess data
-               metric = "RMSE")                    # Use RMSE as the metric
+# Print the RMSE
+print(paste("Null Model RMSE on Test Data:", rmse_null_test))
 
-# Print the cross-validation results
-print(model)
+# Initialize variable to store predictions on the test set
+predictions_lm1_test <- rep(NA, nrow(testData))
+
+# Update the formula to include Age, History, Alcohol, and MTRANS as predictors
+formula_lm1_test <- paste("BMI ~ Age + History + Alcohol + MTRANS")
+
+# Fit the linear regression model using the updated formula and the training data
+model_lm1_test <- lm(formula_lm1_test, data = trainData)
+
+# Make predictions on the test set
+predictions_lm1_test <- predict(model_lm1_test, newdata = testData)
+
+# Calculate RMSE on the test data
+RMSE_lm1_test <- sqrt(mean((testData$BMI - predictions_lm1_test)^2))
+
+# Print the RMSE on the test data
+print(paste("RMSE on test data:", RMSE_lm1_test))
+
+# Define formula for linear model 2
+formula_lm2_test <- as.formula(paste("BMI ~", paste(predictors_updated, collapse = " + ")))
+
+# Fit linear model 2 model using the entire training data
+model_lm2_test <- randomForest(formula = formula_lm2_test, data = trainData)
+
+# Make predictions on the test data
+predictions_lm2_test <- predict(model_lm2_test, newdata = testData)
+
+# Calculate RMSE on test data
+rmse_lm2_test <- sqrt(mean((testData$BMI - predictions_lm2_test)^2))
+
+# Print RMSE on test data
+print(rmse_lm2_test)
+
+# Fit the LASSO model using the optimal lambda
+lasso_model_optimal <- glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, lambda = 0.0011)
+
+# Make predictions on the test data using the optimal LASSO model
+predictions_lasso_test <- predict(lasso_model_optimal, newx = as.matrix(testData[, predictors_updated]))
+
+# Calculate RMSE on the test data
+rmse_lasso_test <- sqrt(mean((testData$BMI - predictions_lasso_test)^2))
+
+# Print RMSE on test data
+print(rmse_lasso_test)
+
+# Create a vector of predicted values with the most frequent category for testData
+predicted_outcome_test <- rep(most_frequent_category, nrow(testData))
+
+# Calculate the accuracy of the null model on testData
+accuracy_null_model_test <- sum(predicted_outcome_test == testData$Obesity)/ length(testData$Obesity)
+
+# Print the accuracy
+print(paste("Accuracy of the null model on test data:", accuracy_null_model_test))
+
+# Define formula for DA model
+formula_DA_test <- as.formula(paste("Obesity ~", paste(predictors_updated, collapse = " + ")))
+
+# Fit the DA model using the entire training data
+model_DA_test <- lda(formula_DA_test, data = trainData)
+
+# Make predictions on the test data
+predicted_DA_test <- predict(model_DA_test, newdata = testData)
+
+# Calculate accuracy on the test data
+accuracy_DA_test <- mean(predicted_DA_test$class == testData$Obesity)
+
+# Print the accuracy
+print(paste("Accuracy of the DA model on test data:", accuracy_DA_test))
 ```

### 2024-04-22 R/eda-code/eda.qmd
part 5

@@
 # histogram of population by age
+# Calculate the bin edges and labels
+bin_data <- mydata %>%
+  mutate(bin = cut(Age, breaks = c(0, 10, 20, 30, 40, 50, 60, 70, Inf), 
+                   labels = FALSE, include.lowest = TRUE)) %>%
+  group_by(bin) %>%
+  summarise(bin_min = min(Age), bin_max = max(Age)) %>%
+  mutate(bin_min = as.integer(bin_min),  # Convert bin_min to integer
+         bin_max = as.integer(bin_max),  # Convert bin_max to integer
+         bin_label = sprintf("%d-%d", bin_min, bin_max))  # Format bin labels
+
 hp2 <- ggplot(mydata, aes(x = Age)) +
-  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
+  geom_histogram(binwidth = 10, fill = "grey", color = "black") +
   labs(x = "Age", y = "Count", title = "Histogram of Age distribution") +
-  stat_bin(binwidth = 5, geom = "text", aes(label = ..count..),
+  stat_bin(binwidth = 10, geom = "text", aes(label = ..count..),
            vjust = -0.5, color = "black", size = 3)  +
-  theme_minimal()
+  geom_text(data = bin_data, aes(x = bin_min + (bin_max - bin_min)/2 - 5, y = -5, label = bin_label),
+            vjust = +1.2, size = 3, color = "black") +  # Add text labels for bins
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 8))  # Adjust the size of x-axis text
 plot(hp2)
+
  # bar chart of obesity level distribution

### 2024-04-22 products/manuscript/Manuscript copy.qmd
part 5

--- 2024-03-23 products/manuscript/Manuscript copy.qmd
@@
 
-In this project, I have found......
+Obesity is a chronic complex disease which may increase the risk of severe health issues including diabetes, cardiovascular disease and even cancers.
+And the condition of obesity in real life could be connected to individual physical behaviors, such as physical activity frequency, alcohol consumption, age, gender and inherit.
+The project is to explore the potential factors that might be related to obesity level using machine learning models to a dataset collected from an investigation done in South American countries including Colombia, Peru and Mexico.
+The data analysis result has shown that age, gender, family history, drinking water, physical activity, and commute transportation are all predictors for obesity level, which is represented by individual BMI.
+As for the outcome, when linear model is fit to data which has BMI as continuous variable, the performance is better than using obesity level as categorical variable.
 
@@
 [Obesity](https://www.who.int/news-room/fact-sheets/detail/obesity-and-overweight)![]() is a chronic complex disease defined by excessive fat deposits that can impair health.
-Obesity can lead to increased risk of type 2 diabetes and heart disease, it can affect bone health and reproduction, it increases the risk of certain cancers.
+Overweight and obesity are predominately a result of a sustained positive energy balance, stemming from a combination of excess dietary energy intake (mainly due to poor eating habits) and reduced energy expenditure (due to lack of physical activity and prolonged sedentary activities).
+It is a significant risk factor for and contributor to increased morbidity and mortality, mainly by its increasing risk of type 2 diabetes and heart disease, as well as certain types of cancer and chronic diseases, including osteoarthritis, liver and kidney disease, sleep apnea, and depression [@pi-sunyer2009].
+Also it does affct bone health and reproduction.
 
-In the recent decades, obesity has become a more and more common and severe problem across the world.
-
+In the recent decades, obesity has become a more and more common and severe problem across the world with worldwide obesity rates tripling since 1975, thus reaching pandemic proportions [@nu15132896].
 Worldwide adult obesity has more than doubled since 1990, and adolescent obesity has quadrupled.
 In 2022, 1 in 8 people in the world were living with obesity.
+In the US.
+the situation seems similar with the worldwide trend.
+In 2022, 22 states had at least 35% of adults with obesity, up from 19 states in 2021.
+Ten years ago, CDC said, no state had an adult obesity prevalence at or above 35%.
 
-There is no doubt that obesity is a multifactorial disease due to obesogenic environments, psycho-social factors and genetic variants in most cases, but personal behavior could also has correlation with the result of overweight and obesity.
+Obesity level is basically defined by Body Mass Index ([BMI](https://www.cdc.gov/obesity/basics/adult-defining.html)).
+BMI is a person’s weight in kilograms divided by the square of height in meters.
+A high BMI can indicate high body fatness.
+BMI less than 18.5 is defined as underweight; BMI range between 18.5-24.9 falls within the healthy weight range.
+If BMI is from 25 to 29.9, then the person is regarded as overweight.
+And when someone's BMI is higher than 30, he/she is at the obesity range.
+
+There is no doubt that obesity is a multifarious disease due to cryogenics environments, psycho-social factors and genetic variants in most cases, but many factors can contribute to excess weight gain including eating patterns, physical activity levels, and sleep routines.
+Social determinants of health, genetics, and taking certain medications also play a role.
 
 There is a lot of discussion on the Internet and social media about fitness and weight loss, and people assume that physical activity can lose weight as a fact.
-Also KOLs ususally encourage people to drink more water for higher metabolic levels.
+Also KOLs usually encourage people to drink more water for higher metabolic levels.
 But I have not seen a clear data to prove it.
@@
 
-In addition, according to what I learned from nutrition course, alcohol consumption may be related to weight gain because [alcohol metabolism](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6527027/) and carbohydrate metabolism both go through the same pathway/cycle in the middle and late stages.
+In addition, according to what I learned from nutrition course, alcohol consumption may be related to weight gain because alcohol metabolism and carbohydrate metabolism both go through the same pathway/cycle in the middle and late stages [@cederbaum2012].
+Ethanol is converted into acetaldehyde by the action of ethanol dehydrogenase, and the acetic acid produced by acetaldehyde dehydrogenase eventually enters the tricarboxylic acid cycle to produce energy, which is consistent with process of the consumption of carbohydrates for energy.
+If someone is lacking of oxidative enzymes, ethanol is nonoxidatively metabolized by two pathways.
+A reaction catalyzed by the enzyme fatty acid ethyl ester (FAEE) synthase leads to the formation of molecules known as FAEEs.
+A reaction with the enzyme phospholipase D (PLD) results in the formation of a phospholipid known as phosphatidyl ethanol [@zakhari2006].
+No matter how calalystic reactions work, finally alcohol will be converted to the reagents for tricarboxylic acid cycle and then energy is released.
 
@@
 The data is obtained from a website [UCI Machine Learning Repository](https://archive.ics.uci.edu/dataset/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition) which have collected the survey results of individuals from the countries of Mexico, Peru and Colombia on their eating habits and physical condition, as well as their estimation of obesity levels.
+The original dataset contains 17 variables and 2111 observations, labeled with the class variable Obesity Level which has 7 categories.
 
@@
 
-Then simple linear fit will be applied to explore the relationship between variables and outcomes (both categorical and numeric).
-
-Prediction will be made.
+Then simple linear fit will be applied to explore the relationship between variables and outcomes (both categorical and numeric), and proper predictors will be pick out from the variables.
+And then machine learning models will be chosen and fitted to the dataset wieh each performance metrics will be determined and compared for reporting the "best" model among the choices based on reality.
 
@@
 
-The dataset presents the results of a survey on the topic of obesity including personal physical conditions and answers to behavioral questions.
+The dataset presents the results of a survey on the topic of obesity including eating habits and personal physical conditions and answers to behavioral questions.
 
@@
 And the code for data cleaning is in the 'processing-code' subfolder under 'R' folder.
+
+I made a code book based on the article reporting this dataset [@PALECHOR2019104344] first before doing any cleaning.
 
@@
 
-By fitting the data with simple linear model, the result will show whether each independent variable has main effect to the dependent variable through the P value.
+By fitting the data with simple linear model, the result will show whether each independent variable has main effect to the dependent variable and the possible interactions between variables will also be detected.
 And the summary table is stored in 'table' subfolder under 'result' folder.
@@
 
-Bar chart shows that the genders of the cohort under investigation are balanced.
+Bar chart shows that the ratio of male and female populaiton is quite close to 1:1, which means genders of the cohort under investigation are balanced.
 
@@
 
-It can be seen that almost two thirds of the people are identified with over weight.
+The age distribution figure shows that people aged 20 to 30 account for the largest proportion of the total population which is nearly two thirds of the total population.
+Then the second highest ranking is 30-40, occupying about a quarter of the total population.
 
 Figure 2.
+Age distribution
+
+![](images/age-distribution-02.png)
+
+It can be seen that almost three fourths of the people are identified with over weight.
+The highest population fall into over weight level 1, while the under weight level has the lowest population.
+
+Figure 3.
 Obesity level distribution.
@@
 
+The gender ratio is close to 1:1 but its distribution under each obesity level is different.
 There is no obvious pattern of obesity levels by gender.
+The ratios of male and female in other obesity levels are close to 1:1 except for overweight level 2 that women occupy almost all of the population and Obese level 1, most of people under which are male.
+In addition, around two thirds of underweight people are men.
 
-Figure 3.
+Figure 4.
 Gender distribution in each obesity levels.
@@
 
-It is apparent that people with family obesity history may tend to be over weight according to figure 4 below.
+Since people aged between 21 and 30 occupy the largest part of the total population, I could expect that in each level, most of people would fall in the age group of 21-30.
+And it can be seen from this bar chart.
+Moreover, more than half of underweight people are from the age group 0-20, it might reflect the problem that health condition and nutrition intake should be regarded seriously by the countries that attend the survey.
 
-Figure 4.
+Figure 5.
+Age distribution in each obesity levels.
+
+![](images/obesity%20level%20by%20age.png)
+
+It is apparent that people with family obesity history may tend to be overweight according to figure 6 below.
+The median BMI is much higher for people who have family obesity history than the people who don't.
+And the data density of no family history group is aggregated at the level under 30.
+There is an article talking about family obesity history and its impact on children in the family.
+The article finds childhood obesity is positively correlated to the obesity status of their parents [@nu15132896].
+When both parents were obese, approximately half of the children in Europe were overweight.
+This highlights the presence of an "obesogenic environment" typically fostered by obese parents, impacting not only themselves but also their children.
+This environment often involves reduced physical activity, increased sedentary behavior such as screen time, and higher intake of calorie-dense foods, contributing to weight gain across the family.
+Therefore, family history is not only a genetic effect, lifestyle, environment also plays an important role for offspring obesity.
+
+Figure 6.
 Violin plot of BMI by family obesity history.
 
-![](images/obesity level by family history-02.png)
+![](images/obesity%20level%20by%20family%20history-01.png)
 
@@
 Also, I use single indicator, alcohol consumption and BMI as an outcome to fit linear model.
+It shows that if an individual consumes alcohol sometimes than never, his BMI increase 3.98.
+But the correlations are all negative for level 3 and 4, which means consuming more alcohol may decrease BMI.
+
+Because I still wonder if obesity could be related to gender or age.
+So I fit simple linear model to the dataset using gender and age as independent variable.
+The result indicates the interactions between them.
+It is a factor out of my expectation.
+And I decide to include this interaction for next machine learning models.
 
@@
 
+Since I think it is quite obvious that family history is an important factor for obesity, and also I have mentioned some reasons how family obesity could influence the kids involved, I choose to skip it for this basic exploratory analysis here.
+But I will include it in the following part.
+
 ## Full analysis
 
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Model                             | Predictor                                 | Metric Value  | Metric Value on testing data |
++===================================+===========================================+===============+==============================+
+| Linear null                       | \-                                        | RMSE=8.00     | RMSE=8.04                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Linear regression 1               | Age, History, Alcohol, MTRANS             | RMSE=0.62     | RMSE=6.61                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Linear regression 2               | All                                       | RMSE=6.30     | RMSE=3.99                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Linear regression 2 + LASSO model | All                                       | RMSE=7.38     | RMSE=7.38                    |
+|                                   |                                           |               |                              |
+|                                   |                                           | l=0.0011      |                              |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Random forest                     | Gender, Age, History, Water, Alcohol, FAF | RMSE=3.22     | RMSE=3.97                    |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Logistic null                     | \-                                        | Accuracy=0.17 | Accuracy=0.15                |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Discriminant analysis             | All                                       | Accuracy=0.50 | Accuracy=0.48                |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+| Multinominal logistic regression  | All                                       | Accuracy=0.58 | Accuracy=0.52                |
++-----------------------------------+-------------------------------------------+---------------+------------------------------+
+
+The above table has listed all of the the models I have chosen to fit the data together with the null models.
+
+When I use BMI as a continuous outcome, linear regression models can be applied to the data.
+Compared to the linear null model, I think all of the ML models perform better according to RMSE value.
+Since I think it is reasonable that physical activity is a potential factor that may influence the obesity level, although linear model with all variables and interactions has higher RMSE, I think regression model 2 that picks all of the predictors is better.
+Then since random forest model selects all predictors and RMSE value is relatively lower among all the linear models, RF model is the choice for linear model.
+
+The figure below shows the prediction condition using random forest model together with the null model on testing data.
+Both sides of the 45 degree line have outliners and the scatters are not clustered close enough to this line which indicates the prediction is not really well.
+
+![](images/Observed-Predicted .png){width="2100"}
+
+As for obesity level as a categorical outcome, logistic models are fit to the data.
+And I use accuracy as the performance metric for this evaluation.
+First, both the models I use have better performance than null model.
+And according to the accuracy, multinominal logistic regression performs better.
+
+Because I have chosen random forest model between linear regression models and multinominal logistic regression between logistic models, I use test data to assess the quality of these two models separately.
+And the new RMSE on test data for random forest model is 3.97, and it is 3.22 when the model is trained and cross validation is applied.
+And the new Accuracy is 0.52 when it is used on test data, also lower than 0.58 when the model is trained.
+The metric value difference is a sign that both of the models are overfitting.
+
+So I try to use test data to assess all of the models in the table above and collect new metric values.
+For linear regression 1, the RMSE on testing data is 6.61, which is much higher than that on training data.
+So this model is not doing prediction well.
+The new RMSE for linear regression 2 is 3.99 lower than 6.30 on training data, which is a sign of good prediction.
+Besides, LASSO model has the same RMSE on both training data and testing data which is 7.38.
+It is also a reasonable result for good prediction.
+Both logistic models have lower accuracy than on test data than training data.
+Based on all the result of the machine learning I have used on dataset, I think linear regression model with LASSO regularization is a good choice among them.
 
@@
 
-*Summarize what you did, what you found and what it means.*
+I chose a dataset that summarize an investigation result done in South American countries about individual physical behaviors and the classification of obesity level based on the information of height and weight obtained from the investigation.
+I had some questions to explore and did data analysis on this dataset.
+
+I found simple linear model fitting can answer my questions, which were that alcohol consumption was correlated to obesity, as well as there were interactions between physical activity frequency and water drinking amount.
+In addition, I found there was nearly no correlation between transportation choice and physical activity, which could answer another question I had.
+At the same time, outside the scope of the problem, I also discovered the interaction between gender and age in this process.
+
+For machine learning model fitting, I found the best predictors for BMI index based on the value of some performance metrics after trying both linear regression and classification model, which were gender, age, family history, water drinking amount, frequency of physical activity, alcohol consumption frequency and commute transporation.
+Although the machine learning models that I used for fitting the dataset are different types and their metrics and not be compared directly, the assessment of model quality using testing data could give a clue for the model selection.
+Considering all the results of performance metrics, I finally chose liner regression model with LASSO regularization with all predictors plus BMI as the outcome.
+The reason was that RMSE calculated of predicted MBI values on testing data split from cleaned dataset was as same as the result when it was trained.
+Also, regularization was applied by LASSO model to reduce overfitting at the same time.
 
@@
 
-*Discuss what you perceive as strengths and limitations of your analysis.*
+I think my analysis has include two different types of model according to the type of outcome and each type include 2 models, which can provide more information and the result can be shown in different aspects as well as more choices were provided.
+Also, interactions between predictors were found and included in the model fitting process.
+Both predictors with and without interactions were all fit and compared with each other.
+And LASSO model was used for decreasing overfitting.
+
+I did tried to do tuning for random forest model but I failed for running it successfully, so this could be a limitation that I have.
+And maybe there could be more metrics that I can also calculate but I only chose 1 for each model.
+Furthermore, chatGPT told me that even if testing result has a sign for well predicting, there still could be other factors for further assessment such as F1-score and so on.
+But base on my understanding, I don't know what other methods I could choose from so I just stopped here.
 

### 2024-04-22 products/manuscript/supplement/Supplementary-Material.qmd
part 5

---
title: "Supplement to Explore the factors that may be associated with obesity"

subtitle: ""
author: Xueyan Hu
date: "`r Sys.Date()`"
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    toc: false
    number-sections: true
    highlight-style: github
    bibliography: ../products/manuscript/dataanalysis-references.bib
csl: ../products/manuscript/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

# Overview

All of the raw data file in csv, processed data in RDS file, descriptive analysis, exploratory analysis,and full data analysis codes and products are included in this project folder.

# Code and file information

-   assets: storing reference pdf files and citation style files

-   data:

    -   raw-data: codebook created by me and original data file downloaded from website

    -   processed-data: cleaned data saved in RDS file

-   products: manuscript in qmd file and its render output with same name in doc file

-   R: including all codes needed

    -   analysis-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file

    -   eda-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file

    -   processing-code: including codes for data cleaning process both in qmd and R script file

-   renv: storage for all of the packages used in this project in R

-   results:

    -   figures: all of the plots in png file created and saved from eda-code and analysis-code

    -   tables: summary tables created and saved from eda-code and 2 linear model tables from analysis-code

{{< pagebreak >}}

# Additional Method Details

I used linear regression model, LASSO model, random forest model for continuous variable BMI as the outcome; and I used discriminant analysis model and multinominal logistic regression for categorical variables obesity level as the outcome.

{{< pagebreak >}}

# Additional results

@tbl-summarytable1 shows a summary of the dataset according to each variable.

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "a dataset summary"
#| echo: FALSE
summarytable1 = readRDS(here("results","tables","summarytable1.rds"))
knitr::kable(summarytable1)
```

@tbl-summarytable2 shows a more detailed summary of the dataset using skim function.

```{r}
#| label: tbl-summarytable2
#| tbl-cap: "a detailed dataset summary"
#| echo: FALSE
summarytable2 = readRDS(here("results","tables","summarytable2.rds"))
knitr::kable(summarytable2)
```

@tbl-lm1table shows linear model fitting result using Water + FAF + Water \* FAF as predictors and BMI as outcome.

```{r}
#| label: tbl-lm1table
#| tbl-cap: "linear model fit 1"
#| echo: FALSE
lm1table = readRDS(here("results","tables","lm1table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm1table)

# Create a data frame from the coefficients
lm1table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm1table_df)
```

@tbl-lm2table shows linear model fitting result using Alcohol as predictors and BMI as outcome.

```{r}
#| label: tbl-lm2table
#| tbl-cap: "linear model fit 2"
#| echo: FALSE
lm2table = readRDS(here("results","tables","lm2table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm2table)

# Create a data frame from the coefficients
lm2table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm2table_df)
```

@tbl-lm3table shows linear model fitting result using Gender + Age + History + Water + Alcohol + FAF + MTRANS as predictors and BMI as outcome.

```{r}
#| label: tbl-lm3table
#| tbl-cap: "linear model fit 3"
#| echo: FALSE
lm3table = readRDS(here("results","tables","lm3table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm3table)

# Create a data frame from the coefficients
lm3table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm3table_df)
```

@tbl-lm4table shows linear model fitting result using Gender \* Age as predictors and BMI as outcome.

```{r}
#| label: tbl-lm4table
#| tbl-cap: "linear model fit 4"
#| echo: FALSE
lm4table = readRDS(here("results","tables","lm4table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm4table)

# Create a data frame from the coefficients
lm4table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm4table_df)
```

{{< pagebreak >}}

# Discussion

For reproducing this project, raw-data folder will be downloaded and stored in a new folder called data and then included in a larger folder with all the codes together. Processing-code qmd file under R folder will be opened and run so the raw data will go through data cleaning process and the cleaned data file will be saved in a folder called process-data under the data folder. Then eda qmd file can be opened and run using processed data file and the outputs will be saved in a newly created fold called results. Results should include tables and figures two subfolders.

Then, exploratory-analysis qmd file under R folder will be involved for the whole data analysis and result will be seen directly in the file.

After all of the qmd files under R folder are run out, the final manuscript will be generated. There are reference related files under assets folders so it will be made sure download and then manuscript qmd file under products fold will be find and run with a doc file saved beside the qmd file.

Supplementary material file is under supplement subfolder and a pdf file will be created and saved after running this qmd file.

{{< pagebreak >}}

# References

### 2024-04-22 R/analysis-code/exploratorycode.R

@@ -48,4 +48,501 @@ lm3_result <- summary(lm3)
 # save the result
 lm_file = here("results", "tables", "lm3table.rds")
 saveRDS(lm3_result, file = lm_file)
- 
+
+
+## ---- lm4 --------
+lm4 <- lm(BMI ~ Gender * Age, data = mydata)
+lm4_result <- summary(lm4)
+
+# save the result
+lm_file = here("results", "tables", "lm4table.rds")
+saveRDS(lm4_result, file = lm_file)
+
+
+## ---- fitting linear models --------
+# Split data into training and testing sets
+set.seed(123)  # Set seed for reproducibility
+trainIndex <- createDataPartition(mydata$BMI, p = 0.8, list = FALSE)
+trainData <- mydata[trainIndex, ]
+testData <- mydata[-trainIndex, ]
+# Create indices for 10-fold cross-validation
+cv_folds <- createFolds(trainData$BMI, k = 10)
+
+# predictors excluding interactions first
+# Define predictors
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+
+# Iterate through all possible combinations of predictors
+for (i in 1:length(predictors)) {
+  subsets1 <- combn(predictors, i)
+  for (j in 1:ncol(subsets1)) {
+    predictors_subset1 <- subsets1[, j]
+    
+    # Initialize variable to store RMSE for current combination of predictors
+    total_rmse_lm1 <- 0
+    
+    # Perform cross-validation
+    for (fold in 1:10) {
+      # Extract indices for training and validation data for current fold
+      train_indices <- unlist(cv_folds[-fold])
+      validation_indices <- cv_folds[[fold]]
+      
+      # Fit model using subset of predictors
+      formula_lm1 <- paste("BMI ~", paste(predictors_subset1, collapse = " + "))
+      model_lm1 <- lm(formula_lm1, data = trainData[train_indices, ])
+      
+      # Make predictions on validation set
+      predictions_lm1 <- predict(model_lm1, newdata = trainData[validation_indices, ])
+      
+      # Calculate RMSE for current fold
+      fold_rmse_lm1 <- sqrt(mean((trainData$BMI[validation_indices] - predictions_lm1)^2))
+      
+      # Add RMSE for current fold to total RMSE
+      total_rmse_lm1 <- total_rmse_lm1 + fold_rmse_lm1
+    }
+    
[…]

### 2024-04-22 R/eda-code/edacode.R

@@ -24,12 +24,26 @@ bc1 <- ggplot(mydata, aes(x = Gender)) +
 plot(bc1)
 
 # histogram of population by age
+# histogram of population by age
+# Calculate the bin edges and labels
+bin_data <- mydata %>%
+  mutate(bin = cut(Age, breaks = c(0, 10, 20, 30, 40, 50, 60, 70, Inf), 
+                   labels = FALSE, include.lowest = TRUE)) %>%
+  group_by(bin) %>%
+  summarise(bin_min = min(Age), bin_max = max(Age)) %>%
+  mutate(bin_min = as.integer(bin_min),  # Convert bin_min to integer
+         bin_max = as.integer(bin_max),  # Convert bin_max to integer
+         bin_label = sprintf("%d-%d", bin_min, bin_max))  # Format bin labels
+
 hp2 <- ggplot(mydata, aes(x = Age)) +
-  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
+  geom_histogram(binwidth = 10, fill = "grey", color = "black") +
   labs(x = "Age", y = "Count", title = "Histogram of Age distribution") +
-  stat_bin(binwidth = 5, geom = "text", aes(label = ..count..),
+  stat_bin(binwidth = 10, geom = "text", aes(label = ..count..),
            vjust = -0.5, color = "black", size = 3)  +
-  theme_minimal()
+  geom_text(data = bin_data, aes(x = bin_min + (bin_max - bin_min)/2 - 5, y = -5, label = bin_label),
+            vjust = +1.2, size = 3, color = "black") +  # Add text labels for bins
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 8))  # Adjust the size of x-axis text
 plot(hp2)
 # bar chart of obesity level distribution
 bc3 <- ggplot(mydata, aes(x = Obesity)) +

### 2024-04-27 hall-project-review.qmd

---
title: Project Review 
author: ELIZABETH HALL
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Explore the factors that may be associated with obesity

Name of project author: Xueyan Hu

Name of project reviewer: Elizabeth Hall

------------------------------------------------------------------------

# Specific project content evaluation

------------------------------------------------------------------------

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The context of the project is well described and comprehensive, detailing definitions, causes, and other risks associated with obesity. Previous and related works are mentioned and cited and I think this project fits well within the context of the existing work. Clarity on why the project was undertaken could be improved (stating what new info this project aims to contribute), but overall the background, context, and motivation are well explained.

### Summary assessment

-   strong contextualization and motivation

------------------------------------------------------------------------

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The hypotheses/questions have their own dedicated section and are clearly stated. How the questions are related to the data is not explicitly stated, and could be explained in further detail.

### Summary assessment

-   question/hypotheses somewhat explained

------------------------------------------------------------------------

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data section in the manuscript is brief, but does provide the source of the data and some information about collection, however it does not provide a description of the data. There is a codebook for the data under `codebook.xlsx` which provides variables, questions, and values.

While the data is explained, I think more detail could be given in the manuscript, and additional information about the data could be added to the codebook. I'd recommend looking at codebook templates/examples on GitHub as I personally found it super helpful to see what others included in theirs. (I ended up using [this](https://gist.github.com/fmaskari/5886dbcc5ab73f2ea663) codebook template).

### Summary assessment

-   source and overall structure of data somewhat explained

------------------------------------------------------------------------

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data cleaning is done well, it is simple but does what it needs to do and is explained within the Quarto files. However there are two additional processing scripts which are not relevant to the project which are present in the `processing-code` folder which need to be removed. There is not much discussion of alternatives, but the exploratory analysis is done well. The exploratory code ran flawlessly, and I was able to generate all graphs. While I think some explanation could be added to the comments of the code, the exploratory analysis otherwise is well explained in the manuscript. There are two additional files in the `eda-code` folder leftover from the template which need to be removed.

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

------------------------------------------------------------------------

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

I think the analysis methods were appropriate for the data. I don't fully understand model fitting in general, but I think that the regression models work well with the data in exploring how various factors influence obesity levels. The different components of the analysis were all done well, and decently explained. My only recommendation is to add more descriptive comments to the code portions.

### Summary assessment

-   strong and reasonable analysis

------------------------------------------------------------------------

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Tables and figures are well done and easy to read and understand. All figures/tables include informative text descriptions and are clearly labeled. I'd recommend numbering your tables/figures (e.g. "Figure 1", "Table 2") to bring them up to publication quality, otherwise they are great!

### Summary assessment

-   results are very well presented

------------------------------------------------------------------------

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

I agreed with the interpretation of the findings, and the strengths and limitations of the study are acknowledged and clearly explained.

The findings of the study are well discussed, but could be stronger. I'd suggest adding a clearer link between your findings and their implications. I'd also suggest comparing these findings with the findings of other studies to see how they compare. This would help bolster your findings, and provide additional context to the results. Otherwise, this part was great!

### Summary assessment

-   minor parts wrong, missing or unclear

------------------------------------------------------------------------

## Further comments

I really like your project so far! My main suggestion is using a couple more sources to compare your findings to, in order to better contextualize the data and bolster your findings. Great job so far! :\]

------------------------------------------------------------------------

# Overall project content evaluation

------------------------------------------------------------------------

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

I think the structure was the weakest part of the project, but this can be easily fixed!

I'd suggest updating all your "read me" files, adding information about the files in each folder, and pointing to additional information where needed. I'd also recommend removing all junk files so that they are not mistaken for project files. If you make an updated codebook, you can also add additional information there about the data processing steps, and what code to run in what order when processing the data to improve clarity.

### Summary assessment

-   poor and confusing structure

------------------------------------------------------------------------

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project is well documented. Code files have corresponding qmd files, which are thorough in explaining the code. My only (minor) gripe is that the actual R files have few comments within the code. I'd suggest updating these, but otherwise the code is very well documented!

### Summary assessment

-   fully and well documented

------------------------------------------------------------------------

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I was unable to reproduce some parts, which I have detailed below:

`exploratorycode.R`

-   Was unable to get the code to work despite manual intervention.

-   I got several errors about missing objects, which I think is due to required packages which are not mentioned in the code.

-   I had to change the data load path to `data_location <- here::here("data","processed-data","processeddata.rds")` in order to get the data to load in.

    -   This issue is fixed in the qmd file, but I'd reccoment updating the main R file as well.

I'd recommend taking a look at your model fitting code and making sure it works. Model fitting is finicky, and I was having similar issues at first and had to fiddle around with things. This is fixable, though! And otherwise things look good, all your other code ran smoothly for me.

### Summary assessment

-   small parts not reproducible or required manual intervention

------------------------------------------------------------------------

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

I think the study was decently thorough. Alternative variables are explored, though alternative machine learning models or analysis methods are not discussed. The questions/hypotheses are fully addressed, though could benefit from some cross referencing with other sources.

My main suggestions would be to discuss a couple alternative analysis methods/ML models and to bolster your findings with the findings from other studies.

### Summary assessment

-   decent level of thoroughness

------------------------------------------------------------------------

## Further comments

Your project is looking good so far! All the issues I listed are easily fixable, and non major. I especially liked how you visualized your data. :\] I can't wait to see this finished!

### 2024-04-29 project-review-RUIZ_files/project-review-RUIZ.qmd
merge files

---
title: Project Review Template 
author: Andrew 
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Factors associated with obesity

Name of project author(s): Hu Xueyan

Name of project reviewer: Andrew Ruiz

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation

Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background provides appropriate amount of information on the topic. This section could benefit from an explanation of how these models might have an impact on incidence, interventions, etc of obesity.

Citations appear to be formated correctly. However, in paragraph 2 of section 2.1, the text mentions stats from the CDC, but there is not a citation for these.

It might also be worth noting, either here or in the limitations, that BMI can be misleading and is not a perfect indicator of obesity.

The term KOL is used in this section, but it is not defined, as far as I can see.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

-   some contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The question is well defined, but it also seems to evolve by the end of the paper. I would recommend re-visiting the question to ensure it covers the scope of the analysis.

### Summary assessment

-   question/hypotheses somewhat explained

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

A code book is available for the data. The source of the data is provided. The overall structure of the data is described.

However, this section could benefit from a better description of the dataset. How were the participants recruited? Was this a random sample? Etc.

### Summary assessment

-   source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data wrangling process is straightforward and well described in the qmd files. It could be better described in the manuscript. Were there any records that were excluded because of missing values, etc.?

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis appears appropriate. However, I am having trouble matching the models in the manuscript table to the models in the qmd. Consider directly referencing the names from the qmd in the manuscript.

### Summary assessment

-   defensible but not optimal analysis

-   <div>

    ## Presentation

    </div>

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The tables and figures were clear. I would consider omitting the bar graph of gender, I don't think it demonstrates the age any better than text.

The table in 4.3 could be improved by reformatting the 'Metric Value' column so that the numbers are not split.

### Summary assessment

-   results are very well presented

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

I think that the reporting of results was appropriate. But the discussion of the results could be improved. What do the findings suggest? How would this inform future research?

The interactions in the models should be discussed with more detail. How did you decide which interactions to include, etc.

Additionally, a better description of the linear model would improve the discussion. Were all levels of the catgorical BMI used?

Strengths and weakness should be expanded to include how the dataset affects the interpretation. Was it a random sample? Does the different countries included in the dataset make generalization easier or harder? What affect might the interaction have on the results? Etc.

### Summary assessment

-   major parts of discussion missing or wrong
-   minor parts wrong, missing or unclear
-   strong, complete and clear discussion

## Further comments

*Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project.*

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

There are some files not in use that are still in the code files. There is also text in some of the used code files left over from the template.

The README files should be updated.

While the structure is uncluttered and not confusing, there is still some clean up to perform before finalizing the project.

### Summary assessment

-   mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)

-   <div>

    ## Documentation

    </div>

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Where the process is documented the objectives are mostly clear. However, the README files need to be updated to guide users on how to use the repository.

One more note: the qmd called exploratory-analysis should be renamed. The name currently implies it should go with the EDA folder

### Summary assessment

-   decently documented with some gaps

-

-   <div>

    ## Reproducibility

    </div>

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The code ran without errors until the final exploratory-analysis file (several chunks beginning at line 155).

Chunk 9 error:

```         
Error: object 'minRMSE1' not found
```

Chunk 11 error:

Error in model.frame.default(Terms, newdata, na.action = na.action, xlev = object\$xlevels) :
factor Alcohol has new levels 4

Chunk 13:

```         
Error in all_predictors[1:i] : 
  object of type 'closure' is not subsettable
```

Chunk 14:

Error in lda.default(x, grouping, ...) :
variable 3 appears to be constant within groups

### Summary assessment

-   small parts not reproducible or required manual intervention

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The study design was straightforward. It was a good decision to calculate a new feature with the actual BMI. Still this category is the least developed in this rubric, but can easily be corrected the some minor adjustments.

The decision to use all seven catgories of BMI for the classification models should have been discussed and alternatives (creating 2 classes obese/not obese, for example) should have been explored.

I think the thoroughness could be improved with expanded discussion, limitations, and data source sections.

### Summary assessment

-   weak level of thoroughness

## Further comments

This was an easy-to-follow, straightforward design. The text might benefit from a review of the perspective and tone. At times it seemed like an academic paper and other times it was more of a narrative.

### 2024-04-29 project-review-template_files/project-review-template.qmd
merge files

---
title: Project Review Template 
author: Taylor Glass
date: 4-22-24
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Explore the factors that may be associated with obesity

Name of project author(s): Xueyan Hu

Name of project reviewer: Taylor Glass

# Specific project content evaluation

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Your background content is thorough, and I appreciate the detailed breakdown of BMI related to obesity. I recommend adding an explanation for what a KOL is and how drinking water can lead to higher metabolic levels. The citations you have included are great, but I think you need to add one for the prevalence information of obesity in the United States and the statements about the factors that affect obesity. Consider adding a source about previous obesity research that focuses on the same factors you study with your dataset.

### Summary assessment
* some contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

I find your questions to be clear, and I know that physical activity and drinking water are the two factors that most closely predict obesity. I know your questions are related to your data because observations are labeled with the obesity class variable that has 7 levels. I think stuying alcohol consumption and obesity will be very interesting.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The source for the data is provided, but I recommend adding details about the website itself and the survey used to collect the information. When I visited the link myself, I see that 77% of the data was generated synthetically, which is worth mentioning. I see there is a variable table on the website as well, but it is great that you also include a codebook in the processing file. However, the codebook you include is only a list of the variables. Is there a way to download the variable table from the website since it has a little more information. I recommend adding a quick list of the variables included in the dataset in your summary to give a better idea of the direction you are headed in with your analysis.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Remember to delete all the information in your quarto documents that is not necessary for your project, including his original instructions! I also think you only need 1 version of the processing file. In your processing file, there is a link to Titanic data, which I do not think is relevant to the project. Your processing code ran smoothly. I loaded the processed data in the EDA document without any issues, and I like that you added blue and pink to the gender bar graph! I think it would be helpful to include an explanation of the 7 different obesity levels because I don't know what the difference between obese1, obese2, obese3, etc is. I think your gender distribution by obesity level graph is really great! The explanation of the violin plot in the manuscript is so thorough, and I am really impressed with the information you included from another source to support your findings. The additional results in the supplementary material are well-labeled and easy to navigate through.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

I would say that the analysis methods are appropriate because linear models would have been my first choice as well, and I think it was smart to investigate the correlation with transportation methods. I think it is great that you tried several combinations of your predictors in your linear models because you got to explore interaction effects. I like that you explained why you used the RMSE performance metric, and I think it is a good choice. I had no issue splitting the data into testing and training pieces, but I got an error saying 'object minRMSE1 not found' when I tried to run iterate through all possible combinations of predictors in your cross-validation for-loop. The next code chunk ran fine, and I see the null model RMSE with cross-validation value. In the next code chunk with the `predictors_updated` object, I get an error saying this:Error in model.frame.default(Terms, newdata, na.action = na.action, xlev = object$xlevels) : factor Alcohol has new levels 4. I think you need to adjust the levels in the new data to match the ones in your training data. I recommend starting by exploring the "new" levels, and then reassigning them to the training data. The LASSO model code ran without any errors, but I did get several warnings about NAs introduced by coercion. I am not sure where they are coming from, but I did some research and found that glmnet does not accept NA values, which can skew your coefficient estimates. I think you need to use the as.numeric() function with your predictors_updated before you run the LASSO regression. If that does not fix it, then I recommend trying to impute the missing values. When I ran the random forest model, I got this error: 'Error in all_predictors[1:i] : object of type 'closure' is not subsettable'. I think the issue might be with your `predictors_updated` object because you include the `predictors` variable in it. I would recommend listing out each predictor in that vector just to be safe. I am super impressed that you attempted discriminant analysis, and I think it is appropriate here because your outcome of BMI is interval based. Unfortunately, I got this error when I ran that code: "Error in lda.default(x, grouping, ...) : variable 3 appears to be constant within groups;" If variable 3 really is constant, then I would consider taking it out of your LDA because it is crucial that there is variability in each term included in the model. If you feel that it is important for your model, then consider creating an interaction term to increase variability. I am wondering if I ran your code out of order, so consider adding instructions in your manuscript about how to reproduce your project if that is the case. The rest of your code ran without any errors from the code chunk creating the null model through the end of the exploratory-analysis.qmd, and I love the table you created with all the different RMSE values before and after testing and training. I can see your observed versus predicted plot, and I agree there is overfitting based on all the scatter. I respect that you did not choose the model with the best RMSE because you feel like physical activity is an important factor for your outcome, and I think choosing the random forest model is a great decision!

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

I think your results are presented well, especially with the explanations you provide from your exploratory data analysis. Violin plots are a little difficult to intepret, but your explanation is great. I find all the other tables and figures to be high quality. I think using `gtsummary` or the `kable()` function to make a table would bring your table up to publication level quality, but it is very clean and easy to read when using the manual quarto syntax. I think it could be helpful to include some tables with coefficient estimates from your linear regression model in the basic statistical analysis section just to provide a little more detail on your linear models. You did a good job acknowledging interaction between some of the predictors, but I think seeing the tables with the coefficient estimates will help your argument. Your explanation of the results from your testing and training data is very helpful and well thought out. You highlight LASSO regression as a good choice here, but I see you picked RF for your final model at this point. I recommend adding a little detail in the full analysis section about why you finally chose the LASSO regression as I saw in your discussion section. I think that would help with consistency.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
I like that you include a full list of predictors that are important to your outcome in your discussion section, and I think you justified why you chose LASSO regression well. I think your strengths and limitations are both appropriate, but I recommend mentioning the overfitting of your model in the limitations section as well. I think the findings for the model with all the predictors are interpreted properly because you acknowledge all the factors that contribute to obesity. I suggest adding information about which predictor was the strongest and weakeast to build out the discussion of your findings because I did not see any information about the LASSO regression coefficients and how much they were reduced from the original model. I also recommend adding a short summary about the gender and age distribution of your dataset that you discussed in your exploratory analysis.

### Summary assessment
* minor parts wrong, missing or unclear

## Further comments

_Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project._
I really like that you explored so many different models and included testing and training models for each one. Since you chose LASSO regression as your final model, I think it would be great to include tuning for your penalty parameter. There is an autoplot() function that will show you the RMSE scores based on parameter penalties once you use the tuning grid like we did in the class exercise, which will help you determine if you used an appropriate penalty. Other than that, I feel like you explored many things with a lot of detail!

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is well-structured, and all the files have reasonable names. I like that you grouped all the code together, but I recommend changing the "R" folder name to "code" because the whole project is in R. I get an idea of where to start by just looking at the files and folders, but some of the junk files still need to be removed from the 3 code folders. I also see a couple copies of the manuscript that I do not think need to be kept. Make sure to update all the READMe files as well!

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

I recommend updating the main README.md file to include information on how to reproduce your project. I see that your supplementary material includes where to start for reproducing the project, but I do not think that is the best location for this information. I think more information in the methods section would be helpful. Consider adding a full list of the models you chose with an explanation of what the performance metrics mean. I think the data cleaning section in your processing file could include more comments in the code because I am not completely sure why you recoded the gender and alcohol variables the way you did. I think your eda and exploratory-analysis codes are well explained with a lot of comments. I understand why you made each decision, so just consider adding details about why you chose the LASSO regression for the final model as I discussed earlier.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I had trouble reproducing the exploratory-analysis.qmd with the cross validation techniques. These are the errors I encountered:
1) Error: object 'minRMSE1' not found
2) Error in model.frame.default(Terms, newdata, na.action = na.action, xlev = object$xlevels) : 
  factor Alcohol has new levels 4
3) Error in all_predictors[1:i] : 
  object of type 'closure' is not subsettable
4) Error in lda.default(x, grouping, ...) :
variable 3 appears to be constant within groups
I did not know how to manually intervene or edit in these 4 scenarios. I didn't want to make major changes to your code without consulting you, so I added suggestions for how to fix these errors earlier. All of the other files were fully reproducible without any manual intervention, and I was able to generate all of your figures and tables.

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

I think your overall study is very thorough considering you explored a wide range of the models we discussed in class! You covered all the alternative options well, and I like that you created the table comparing performance metrics with training and testing data for each model. I think you did a great job discussing alternatives and explaining why you made your final decisions. Your questions were thoroughly addressed considering each factor affecting obesity was included in your model. Consider trying the tuning process for the random forest model again if you have time based on our in class exercise! Tuning is very important in machine learning, but I think that would be the only way to improve the thoroughness of your project.

### Summary assessment
* strong level of thorougness

## Further comments

_Add any other comments regarding the overall project here. Write anything you think can help your classmate improve their project._
I really enjoyed exploring your project, and I am very impressed with all the machine learning techniques you used! I think including how to reproduce your project in the methods section of your manuscript would be very helpful, but that is my last sugggestion. You did an amazing job!

### 2024-05-03 Rcode/dneda-code/dneda.qmd
improve after review

--- 2024-04-22 R/eda-code/eda.qmd
@@
 
-This Quarto file loads the cleaned data and does some exploring.
-
-I'm only showing it the way where the code is included in the file. 
-As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.
-
-But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)
-
-Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
-And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.
-
-Also note that while here I split cleaning and exploring, this is iterative.
-You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.
-In general, as you explore, you'll find things that need cleaning.
-As you clean, you can explore more.
-Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file.
-Or split things in any other logical way.
-
-As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-Figures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data.
-
-Start by exploring one variable at a time.
-Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-
-Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.
+This file loads the processed data cleaned from processingfile-v1 and basic descriptive analysis is applied in this file.
 

### 2024-05-03 Rcode/mlmodel-code/ml-model-v1.qmd
improve after review

--- 2024-04-22 R/analysis-code/exploratory-analysis.qmd
@@
 ---
-
-This Quarto file loads the cleaned data and does some exploring.
-
-I'm only showing it the way where the code is included in the file.
-As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.
-
-But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)
-
-Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
-And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.
-
-Also note that while here I split cleaning and exploring, this is iterative.
-You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.
-In general, as you explore, you'll find things that need cleaning.
-As you clean, you can explore more.
-Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file.
-Or split things in any other logical way.
-
-As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-Figures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data.
-
-Start by exploring one variable at a time.
-Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-
-Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.
 
@@
 ```{r}
+# Filter out observations where Alcohol is 4
+mydata1 <- mydata[mydata$Alcohol != 4, ]
 # Split data into training and testing sets
@@
 trainIndex <- createDataPartition(mydata$BMI, p = 0.8, list = FALSE)
-trainData <- mydata[trainIndex, ]
-testData <- mydata[-trainIndex, ]
+trainData <- mydata1[trainIndex, ]
+testData <- mydata1[-trainIndex, ]
 # Create indices for 10-fold cross-validation
@@
 ```
+
+Since there is only one observation with alcohol consumption level at 4, so after splitting the data, it will be distributed randomly in either trainData or testData.
+When it is in trainData and the cross-validation is applied, there will be some folds doesn't have this level and error will occur during this process.
+So I drop it from the dataset to make it working easier.
+
+I will fit the first linear model which is called linear model 1 to the data and use cross-validation, then calculate RMSE.
+There is no interactions in this model's predictors.
+And by running this chunk, predictor combination that indicates the lowest RMSE will be selected and shown in the output together with its RMSE.
 
@@
 predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+
+# Set seed
+set.seed(123)
 
@@
 
+Then I will calculate RMSE for null model of linear regression.
+
 ```{r}
@@
 
+Using the same linear regression named linear model 2 but adding 2 interactions as predictors and under these predictors, best combination with lowest RMSE will be shown in the output.
+
 ```{r}
@@
 bestSubset2 <- NULL
+
+# Set seed
+set.seed(123)
 
@@
 ```{r}
+# Create dummy variables for categorical variables in trainData
+trainData1 <- within(trainData, {
+  History <- as.factor(History)
+  MTRANS <- as.factor(MTRANS)
+})
+dummy_vars_train <- model.matrix(~ History + MTRANS - 1, data = trainData1)
+
+# Combine dummy variables with original trainData
+trainData1 <- cbind(trainData, dummy_vars_train)
+
+# Repeat the same process for testData
+testData1 <- within(testData, {
+  History <- as.factor(History)
+  MTRANS <- as.factor(MTRANS)
+})
+dummy_vars_test <- model.matrix(~ History + MTRANS - 1, data = testData1)
+testData1 <- cbind(testData, dummy_vars_test)
+
+# Define predictors and outcome
+response <- "BMI"
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+interaction_terms <- c("Gender_Age", "FAF_Water")
+
+# Update predictors to include dummy variables
+predictors_lasso <- c("Gender", "Age", "Water", "Alcohol", "FAF","Gender_Age", "FAF_Water", colnames(dummy_vars_train))
+
+# Define function to calculate RMSE
+calculate_rmse <- function(predictions, actual) {
+  return(sqrt(mean((actual - predictions)^2)))
+}
+
 # Initialize variables to store results
@@
 
+# Set seed
+set.seed(123)
+
 # Perform LASSO regression with cross-validation
-lasso_model <- cv.glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, nfolds = 10)
+for (fold in 1:10) {
+  # Extract indices for training and validation data for current fold
+  train_indices <- cv_folds[[fold]]
+  validation_indices <- setdiff(1:nrow(trainData1), train_indices)
+
+  # Define training and validation data subsets
+  train_data_fold <- trainData1[train_indices, ]
+  validation_data_fold <- trainData1[validation_indices, ]
+
+  # Create design matrix for training and validation data
+  train_matrix <- model.matrix(~., data = train_data_fold[, predictors_lasso])[,-1]
+  validation_matrix <- model.matrix(~., data = validation_data_fold[, predictors_lasso])[,-1]
+
+  # Ensure the same variables are present in training and validation matrices
+  common_variables <- intersect(colnames(train_matrix), colnames(validation_matrix))
+  train_matrix <- train_matrix[, common_variables]
+  validation_matrix <- validation_matrix[, common_variables]
+
+  # Perform LASSO regression with cross-validation
+  lasso_model <- cv.glmnet(train_matrix, train_data_fold$BMI, alpha = 1, lambda = 10^seq(10, -2, length = 100))
+
+  # Extract optimal lambda value
+  optimal_lambda <- lasso_model$lambda.min
+
+  # Fit LASSO model using optimal lambda on training data
+  lasso_model_optimal <- glmnet(train_matrix, train_data_fold$BMI, alpha = 1, lambda = optimal_lambda)
+
+  # Make predictions on validation set
+  predictions_lasso <- predict(lasso_model_optimal, newx = validation_matrix, s = optimal_lambda)
+
+  # Calculate RMSE on validation set
+  rmse <- sqrt(mean((validation_data_fold$BMI - predictions_lasso)^2))
+
+  # Update min_rmse and best_subset if CV RMSE is lower
+  if (rmse < min_rmse_lasso) {
+    min_rmse_lasso <- rmse
+    best_subset_lasso <- predictors_lasso  # Include interaction terms
+  }
+}
+
+# Extract non-zero coefficients from the optimal LASSO model
+non_zero_coeffs <- coef(lasso_model_optimal)
+non_zero_coeffs <- non_zero_coeffs[-1]  # Remove intercept
+
+# Find the indices of non-zero coefficients
+non_zero_indices <- which(non_zero_coeffs != 0)
+
+# Extract the names of predictors with non-zero coefficients
+best_subset_names_lasso <- colnames(train_matrix)[non_zero_indices]
 
@@
 
-# Fit the LASSO model using the optimal lambda
-lasso_model_optimal <- glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, lambda = optimal_lambda)
-
-# Compute RMSE on test data using the optimal LASSO model
-predictions_lasso <- predict(lasso_model_optimal, newx = as.matrix(testData[, predictors_updated]))
-rmse_lasso <- sqrt(mean((testData$BMI - predictions_lasso)^2))
-
-# Extract coefficients of the LASSO model
-lasso_coefficients <- as.matrix(coef(lasso_model_optimal))
-
 # Print results
+cat("Lowest RMSE (LASSO CV):", min_rmse_lasso, "\n")
+cat("Best subset of predictors:", paste(best_subset_names_lasso, collapse = ", "), "\n")
+# Print optimal lambda value
 cat("Optimal lambda:", optimal_lambda, "\n")
-cat("RMSE with LASSO:", rmse_lasso, "\n")
-
-# Print non-zero coefficients (selected predictors)
-selected_predictors <- rownames(lasso_coefficients)[lasso_coefficients != 0, drop = FALSE]
-print("Selected predictors by LASSO:")
-print(selected_predictors)
 ```
 
-After that, I will select another regression model to fit data, here I will use random forest model.
+Since I already get optimal lambda above, so I will skip using autoplot function creating figure and tuning the LASSO model.
+
+For trying with more models as comparison, I select another regression model to fit data, here I will use random forest model.
 
@@
 best_predictors_rf <- NULL
+
+# Set seed
+set.seed(123)
 
@@
   # Forward selection
-  for (i in seq_along(all_predictors)) {
-    predictors_subset_rf <- all_predictors[1:i]
+  for (i in seq_along(predictors_updated)) {
+    predictors_subset_rf <- predictors_updated[1:i]
     formula_subset_rf <- as.formula(paste(response, "~", paste(predictors_subset_rf, collapse = " + ")))
@@
     # Fit random forest model using the training data for the current fold
-    model_rf <- randomForest(formula = formula_subset_rf, data = mydata[train_indices, ])
+    model_rf <- randomForest(formula = formula_subset_rf, data = trainData[train_indices, ])
 
     # Make predictions on the validation data for the current fold
-    predictions_rf <- predict(model_rf, newdata = mydata[validation_indices, ])
+    predictions_rf <- predict(model_rf, newdata = trainData[validation_indices, ])
 
     # Calculate RMSE for the current fold
-    rmse_rf <- sqrt(mean((mydata$BMI[validation_indices] - predictions_rf)^2))
+    rmse_rf <- sqrt(mean((trainData$BMI[validation_indices] - predictions_rf)^2))
 
@@
 
+# Set seed
+set.seed(123)
+
 # Create cross-validation folds
@@
       formula_DA <- as.formula(paste(outcome, "~", paste(predictors_subset_DA, collapse = " + ")))
-      model_DA <- lda(formula_DA, data = mydata[train_indices, ])
+      model_DA <- lda(formula_DA, data = trainData[train_indices, ])
 
@@
       # Calculate accuracy for the current fold
-      fold_accuracies[fold_index] <- mean(predicted_DA$class == mydata$Obesity[validation_indices])
+      fold_accuracies[fold_index] <- mean(predicted_DA$class == trainData$Obesity[validation_indices])
     }
@@
 
+# Set seed
+set.seed(123)
+
 # Perform cross-validation
@@
 
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
 | Model                             | Predictor                                                     | Metric Value  | Metric Value on testing data |
-+===================================+===========================================+===============+==============================+
-| Linear null                       | \-                                        | RMSE=8.00     | RMSE=8.04                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Linear regression 1               | Age, History, Alcohol, MTRANS             | RMSE=0.62     | RMSE=6.61                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Linear regression 2               | All                                       | RMSE=6.30     | RMSE=3.99                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Linear regression 2 + LASSO model | All                                       | RMSE=7.38     | RMSE=7.38                    |
++===================================+===============================================================+===============+==============================+
+| Linear null                       | \-                                                            | RMSE=8.01     | RMSE=7.99                    |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Linear regression 1               | Age, History, Alcohol, MTRANS                                 | RMSE=0.62     | RMSE=6.54                    |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Linear regression 2               | All                                                           | RMSE=6.29     | RMSE=3.81                    |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Linear regression 2 + LASSO model | All                                                           | RMSE=6.34     | RMSE=6.47                    |
 |                                   |                                                               |               |                              |
-|                                   |                                           | l=0.0011      |                              |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Random forest                     | Gender, Age, History, Water, Alcohol, FAF | RMSE=3.22     | RMSE=3.97                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Logistic null                     | \-                                        | Accuracy=0.17 |                              |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Discriminant analysis             | All                                       | Accuracy=0.50 | Accuracy=0.48                |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Multinominal logistic regression  | All                                       | Accuracy=0.58 | Accuracy=0.52                |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
+|                                   |                                                               | l=0.023       |                              |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Random forest                     | All                                                           | RMSE=3.35     | RMSE=3.97                    |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Logistic null                     | \-                                                            | Accuracy=0.17 | Accuracy=0.15                |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Discriminant analysis             | Gender, Age, History, Water, Alcohol, FAF, MTRANS, Gender_Age | Accuracy=0.50 | Accuracy=0.50                |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Multinominal logistic regression  | All                                                           | Accuracy=0.55 | Accuracy=0.50                |
++-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
 
-Then I will use testData to assess the quality to the model that I choose.
+Then I will use testData to assess the quality to the model that I choose. Between 2 linear regression models I choose random forest model.
 
@@
 
+Between 2 logistic models, I choose mlr model.
+
 ```{r}
@@
 
-Here I try to assess every model I choose and check the result.
+Since the neither of the model above has satisfying performance, I will try to assess every model I have used for analysis and check the result.
 
@@
 # Fit the LASSO model using the optimal lambda
-lasso_model_optimal <- glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, lambda = 0.0011)
+lasso_model_optimal <- glmnet(as.matrix(trainData1[, predictors_lasso]), trainData1$BMI, alpha = 1, lambda = 0.023)
 
 # Make predictions on the test data using the optimal LASSO model
-predictions_lasso_test <- predict(lasso_model_optimal, newx = as.matrix(testData[, predictors_updated]))
+predictions_lasso_test <- predict(lasso_model_optimal, newx = as.matrix(testData1[, predictors_lasso]))
 
 # Calculate RMSE on the test data
-rmse_lasso_test <- sqrt(mean((testData$BMI - predictions_lasso_test)^2))
+rmse_lasso_test <- sqrt(mean((testData1$BMI - predictions_lasso_test)^2))
 
@@
 
+# Create new predictors for DA model
+predictors_DA <- c("Gender", "Age", "History",  "Water",  "Alcohol", "FAF",  "MTRANS", "Gender_Age")
+
 # Define formula for DA model
-formula_DA_test <- as.formula(paste("Obesity ~", paste(predictors_updated, collapse = " + ")))
+formula_DA_test <- as.formula(paste("Obesity ~ ", paste(predictors_DA, collapse = " + ")))
 
@@
 ```
+I will explore more information for LASSO model.
+```{r}
+# Extract coefficients from the fitted LASSO model
+lasso_coefficients <- coef(lasso_model_optimal)
 
-# Notes
+# Extract non-zero coefficients and their corresponding predictors
+non_zero_indices <- which(lasso_coefficients != 0, arr.ind = TRUE)
+non_zero_coefficients <- lasso_coefficients[non_zero_indices]
 
-For your own explorations, tables and figures can be "quick and dirty".
-As long as you can see what's going on, there is no need to polish them.
-That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.).
-Those should look as nice, polished and easy to understand as possible.
+# Extract names of predictors with non-zero coefficients
+strongest_predictors <- rownames(lasso_coefficients)[non_zero_indices[, "col"]]
+
+# Find the index of the strongest and weakest coefficients
+strongest_index <- which.max(abs(non_zero_coefficients))
+weakest_index <- which.min(abs(non_zero_coefficients))
+
+# Extract the strongest and weakest coefficients and their predictors
+strongest_coefficient <- non_zero_coefficients[strongest_index]
+weakest_coefficient <- non_zero_coefficients[weakest_index]
+strongest_predictor <- strongest_predictors[strongest_index]
+weakest_predictor <- strongest_predictors[weakest_index]
+
+# Print the results
+cat("Strongest predictor:", strongest_predictor, "\n")
+cat("Strongest coefficient:", strongest_coefficient, "\n")
+cat("Weakest predictor:", weakest_predictor, "\n")
+cat("Weakest coefficient:", weakest_coefficient, "\n")
+```

### 2024-05-03 Rcode/mlmodel-code/ml-model-v2.qmd
improve after review

--- 2024-02-24 R/analysis-code/exploratory-analysis-v2.qmd
@@
 
-This Quarto file loads the cleaned data and does some exploring.
+This Quarto file has the content of machine learning model analysis.
 
-This is essentially the same as the other `exploratory_analysis` Quarto file, but now the code is not inside this file.
-Instead, it is pulled in from the R script `exploratorycode.R` using the code chunk labels.
+This is essentially the same as the other `ml-model-v1` Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script `ml-model.R` using the code chunk labels.
 
@@
 
-```{r,table 1}
+```{r,logistic}
 ```
@@
 
-```{r,ANOVA table}
+## lm1
+
+```{r,lm1}
 ```
 
-# LSD test
+## lm2
 
-```{r,LSD test}
+```{r,lm2}
 ```
+
+## lm3
+
+```{r,lm3}
+```
+
+## lm4
+
+```{r,lm4}
+```
+
+## model training
+
+```{r,fitting linear models}
+```
+
+```{r,fitting logistic models}
+```
+
+## model assessment using testing data
+
+```{r,test models}
+```
+
 # Notes

### 2024-05-03 Rcode/processing-code/processingfile-v1.qmd
improve after review

--- 2024-03-23 R/processing-code/processingfile-v1.qmd
@@
 # Processing script
-
-This contains the same code and comments/information as `processingcode.R`.
-
-This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.
-
-See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.
 
@@
 ```
+
 Several ways of looking at the data
-
-The meaning of each variable can be checked [online](https://www.kaggle.com/datasets/yasserh/titanic-dataset/data).
 
@@
 First generating a new variable BMI using height and weight, as well as change some variable names and categorical levels.
+For more intuitive observation and writing, I will change gender categories to 1 and 2 and name alcohol consumption levels to from 1 to 4.
 
@@
 
-Finally, we save the clean data as RDS file.
-I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
-This preserves coding like factors, characters, numeric, etc.
-If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text.
-If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
-
 ```{r}
@@
 ```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g.
-keep individuals with some missing information).

### 2024-05-03 products/manuscript/Manuscript copy.qmd
improve after review

@@
 It is a significant risk factor for and contributor to increased morbidity and mortality, mainly by its increasing risk of type 2 diabetes and heart disease, as well as certain types of cancer and chronic diseases, including osteoarthritis, liver and kidney disease, sleep apnea, and depression [@pi-sunyer2009].
-Also it does affct bone health and reproduction.
+Also it does affect bone health and reproduction.
 
@@
 In 2022, 22 states had at least 35% of adults with obesity, up from 19 states in 2021.
-Ten years ago, CDC said, no state had an adult obesity prevalence at or above 35%.
+Ten years ago, according to [CNN Health](https://www.cnn.com/2023/09/21/health/obesity-more-common-states-cdc-data/index.html), CDC said there was no state had an adult obesity prevalence at or above 35%.
 
 Obesity level is basically defined by Body Mass Index ([BMI](https://www.cdc.gov/obesity/basics/adult-defining.html)).
-BMI is a person’s weight in kilograms divided by the square of height in meters.
+BMI is a person's weight in kilograms divided by the square of height in meters.
 A high BMI can indicate high body fatness.
@@
 Social determinants of health, genetics, and taking certain medications also play a role.
+There is an [article](https://www.nichd.nih.gov/health/topics/obesity/conditioninfo/cause) by NIH listing some causative factors for obesity, including food and activity, environment effect, genetics, health conditions and medications, stress, emotional factors, and poor sleep.
+Other [articles](https://ro.co/weight-loss/obesity-causes-risk-factors-strategies/#risk-factors-for-obesity) also have mentioned hormone as cause and family history, age, sedentary lifestyle and smoking as risk factors for obesity.
+But alcohol consumption is not that frequently mentioned as far as I know.
 
 There is a lot of discussion on the Internet and social media about fitness and weight loss, and people assume that physical activity can lose weight as a fact.
-Also KOLs usually encourage people to drink more water for higher metabolic levels.
-But I have not seen a clear data to prove it.
+Also KOLs (Key Opinion Leader) usually encourage people to drink more water for higher metabolic levels.
+Although I haven't seen any actual data for proving it, some papers have pointed the effect of drinking a lot of water on body weight [@vinu2013].
 At the same time, the effect of physical activity to lose weight is also different from person to person.
@@
 
-The data is obtained from a website [UCI Machine Learning Repository](https://archive.ics.uci.edu/dataset/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition) which have collected the survey results of individuals from the countries of Mexico, Peru and Colombia on their eating habits and physical condition, as well as their estimation of obesity levels.
-The original dataset contains 17 variables and 2111 observations, labeled with the class variable Obesity Level which has 7 categories.
+The data is obtained from a website [UCI Machine Learning Repository](https://archive.ics.uci.edu/dataset/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition) together with a paper as a source, which have collected the survey results of individuals from the countries of Mexico, Peru and Colombia on their eating habits and physical condition, as well as their estimation of obesity levels.
+The original dataset contains 17 variables and 2111 observations, labeled with the class variable Obesity Level which has 7 categories [@PALECHOR2019104344].
+The variables include gender, age, height, weight, family history, eating habits (high caloric food, vegetables, main meals, snacks), smoking, alcohol consumption, water consumption, physical activity, technological devices usage, commute transportation, and their obesity level according to the BMI of each observation but BMI value is not in the dataset directly.
+Over weight level 1 refers to people whose BMI is in the range of 25-27.5, while weight level 2 is 27.5-30.
+But these 2 levels couldn't be found on CDC website.
+Obesity is frequently subdivided into categories: Class 1: BMI of 30 to \< 3; Class 2: BMI of 35 to \< 40; Class 3: BMI of 40 or higher.
 
@@
 
-I would like to figure out if higher frequency of physical activity and drinking more water could keep people away from obesity and whether heredity is an important factor for obesity.
-Also, I will focus on the relationship between alcohol consumption frequency and obesity level.
+Several questions will be explored during the following analysis.
+
+1.  If higher frequency of physical activity and drinking more water could keep people away from obesity.
+2.  Whether there is a relationship between alcohol consumption frequency and obesity level.
+3.  What variables/risk factors could be good predictors for obesity level, and what machine learning model could be fit and make good predictions for new given data.
 
@@
 
-Then simple linear fit will be applied to explore the relationship between variables and outcomes (both categorical and numeric), and proper predictors will be pick out from the variables.
-And then machine learning models will be chosen and fitted to the dataset wieh each performance metrics will be determined and compared for reporting the "best" model among the choices based on reality.
+Then simple linear model will be applied to explore the relationship between variables and outcomes (both categorical and numeric), and proper predictors will be pick out from the variables.
+And then machine learning models will be chosen and fitted to the dataset and each performance metric will be determined and compared for reporting the "best" model among the choices based on reality.
 
@@
 The dataset presents the results of a survey on the topic of obesity including eating habits and personal physical conditions and answers to behavioral questions.
+The paper have pointed out that 23% of data are collected by themselves directly from users through a web platform while 77% of the data is synthesized using the Weka tool and the SMOTE filter.
 
@@
 I made a code book based on the article reporting this dataset [@PALECHOR2019104344] first before doing any cleaning.
+
+There is no missing value in this dataset.
 
@@
 
+![](images/gender-distribution.png)
+
 Figure 1.
 Gender distribution.
-
-![](images/gender-distribution.png)
 
@@
 
+![](images/age-distribution-02.png)
+
 Figure 2.
 Age distribution
-
-![](images/age-distribution-02.png)
 
@@
 
+![](images/obesity-distribution.png)
+
 Figure 3.
 Obesity level distribution.
-
-![](images/obesity-distribution.png)
 
@@
 
+![](images/obesity%20level%20by%20gender.png)
+
 Figure 4.
 Gender distribution in each obesity levels.
-
-![](images/obesity%20level%20by%20gender.png)
 
@@
 
+![](images/obesity%20level%20by%20age.png)
+
 Figure 5.
 Age distribution in each obesity levels.
-
-![](images/obesity%20level%20by%20age.png)
 
@@
 
+![](images/obesity%20level%20by%20family%20history-01.png)
+
 Figure 6.
 Violin plot of BMI by family obesity history.
-
-![](images/obesity%20level%20by%20family%20history-01.png)
 
@@
 
-First, I was curious that whether there is a negative correlation between transportation type and physical activity frequency, I converted transportation types into numbers indicating the energy consuming levels from 1 (lowest automobile) to 4 (highest walking).
-Then I used correlation function to calculate the coefficient for these two variable but it was pretty low (0.0059).
+First, for exploring the correlation between transportation type and physical activity frequency, transportation types were converted into numbers indicating the energy consuming levels from 1 (lowest automobile) to 4 (highest walking).
+Then correlation function was used to calculate the coefficient for these two variable but it was pretty low (0.0059).
 
-Then, I would like to fit linear model for BMI by both water consumption and physical activity.
+Then, linear model for BMI is fit by both water consumption and physical activity.
 And the result indicates that there is a connection effect between them two, which means that both of them contribution to obesity control.
 
-Also, I use single indicator, alcohol consumption and BMI as an outcome to fit linear model.
+Also, single indicator, alcohol consumption is used and BMI as an outcome to fit linear model.
 It shows that if an individual consumes alcohol sometimes than never, his BMI increase 3.98.
@@
 
-Because I still wonder if obesity could be related to gender or age.
-So I fit simple linear model to the dataset using gender and age as independent variable.
+Although age distribution is not so balanced in the dataset, the linear model of obesity by gender and age is still tired to explored.
+So simple linear model is fit to the dataset using gender and age as independent variables.
 The result indicates the interactions between them.
 It is a factor out of my expectation.
-And I decide to include this interaction for next machine learning models.
+And this interaction is considered to be included for next machine learning models.
 
-In the last model I select all the variables that I think may have effect on BMI as predictors and BMI as outcome.
+In the last model selected is all the variables that I think may have effect on BMI as predictors and BMI as outcome.
 
-Since I think it is quite obvious that family history is an important factor for obesity, and also I have mentioned some reasons how family obesity could influence the kids involved, I choose to skip it for this basic exploratory analysis here.
-But I will include it in the following part.
+Since it is quite obvious that family history is an important factor for obesity, and also some reasons how family obesity could influence the kids involved have been mentioned above in the introduction part, it is skipped for this basic exploratory analysis here, but will be included in the following part.
 
@@
 
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
+Table 1.
+Summary of each machine learning model applied to the data and its metric value comparison.
+
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
 | Model                            | Predictor                                                     | Metric Value  | Metric Value on testing data |
-+===================================+===========================================+===============+==============================+
-| Linear null                       | \-                                        | RMSE=8.00     | RMSE=8.04                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Linear regression 1               | Age, History, Alcohol, MTRANS             | RMSE=0.62     | RMSE=6.61                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Linear regression 2               | All                                       | RMSE=6.30     | RMSE=3.99                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Linear regression 2 + LASSO model | All                                       | RMSE=7.38     | RMSE=7.38                    |
++==================================+===============================================================+===============+==============================+
+| Linear null                      | \-                                                            | RMSE=8.01     | RMSE=7.99                    |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Linear regression 1              | Age, History, Alcohol, MTRANS                                 | RMSE=0.62     | RMSE=6.54                    |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Linear regression 2              | All                                                           | RMSE=6.29     | RMSE=3.81                    |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| LASSO model                      | All                                                           | RMSE=6.34     | RMSE=6.47                    |
 |                                  |                                                               |               |                              |
-|                                   |                                           | l=0.0011      |                              |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Random forest                     | Gender, Age, History, Water, Alcohol, FAF | RMSE=3.22     | RMSE=3.97                    |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
+|                                  |                                                               | l=0.023       |                              |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Random forest                    | All                                                           | RMSE=3.35     | RMSE=3.97                    |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
 | Logistic null                    | \-                                                            | Accuracy=0.17 | Accuracy=0.15                |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Discriminant analysis             | All                                       | Accuracy=0.50 | Accuracy=0.48                |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
-| Multinominal logistic regression  | All                                       | Accuracy=0.58 | Accuracy=0.52                |
-+-----------------------------------+-------------------------------------------+---------------+------------------------------+
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Discriminant analysis            | Gender, Age, History, Water, Alcohol, FAF, MTRANS, Gender_Age | Accuracy=0.50 | Accuracy=0.50                |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
+| Multinominal logistic regression | All                                                           | Accuracy=0.55 | Accuracy=0.50                |
++----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
 
-The above table has listed all of the the models I have chosen to fit the data together with the null models.
+The above table has listed all of the the models have been chosen to fit the data together with the null models.
 
-When I use BMI as a continuous outcome, linear regression models can be applied to the data.
-Compared to the linear null model, I think all of the ML models perform better according to RMSE value.
-Since I think it is reasonable that physical activity is a potential factor that may influence the obesity level, although linear model with all variables and interactions has higher RMSE, I think regression model 2 that picks all of the predictors is better.
+When BMI is used as a continuous outcome, linear regression models can be applied to the data.
+Compared to the linear null model, all of the ML models perform better according to RMSE value.
+Since it is reasonable that physical activity is a potential factor that may influence the obesity level, although linear model with all variables and interactions has higher RMSE, regression model 2 that picks all of the predictors is better.
 Then since random forest model selects all predictors and RMSE value is relatively lower among all the linear models, RF model is the choice for linear model.
@@
 
-![](images/Observed-Predicted .png){width="2100"}
+![](images/Observed-Predicted%20.png){width="2100"}
 
 As for obesity level as a categorical outcome, logistic models are fit to the data.
-And I use accuracy as the performance metric for this evaluation.
-First, both the models I use have better performance than null model.
+And accuracy is used as the performance metric for this evaluation.
+First, both the models have better performance than null model.
 And according to the accuracy, multinominal logistic regression performs better.
 
-Because I have chosen random forest model between linear regression models and multinominal logistic regression between logistic models, I use test data to assess the quality of these two models separately.
-And the new RMSE on test data for random forest model is 3.97, and it is 3.22 when the model is trained and cross validation is applied.
-And the new Accuracy is 0.52 when it is used on test data, also lower than 0.58 when the model is trained.
+Because random forest model is chosen between linear regression models and multinominal logistic regression between logistic models, test data is used to assess the quality of these two models separately.
+And the new RMSE on test data for random forest model is 3.97, and it is 3.35 when the model is trained and cross validation is applied.
+And the new accuracy is 0.50 when it is used on test data which is the same as it is trained.
 The metric value difference is a sign that both of the models are overfitting.
 
-So I try to use test data to assess all of the models in the table above and collect new metric values.
-For linear regression 1, the RMSE on testing data is 6.61, which is much higher than that on training data.
+So all of the models are assessed by making predictions test data in the table above and new metric values are collected again.
+For linear regression 1, the RMSE on testing data is 6.54, which is much higher than that on training data.
 So this model is not doing prediction well.
-The new RMSE for linear regression 2 is 3.99 lower than 6.30 on training data, which is a sign of good prediction.
-Besides, LASSO model has the same RMSE on both training data and testing data which is 7.38.
-It is also a reasonable result for good prediction.
-Both logistic models have lower accuracy than on test data than training data.
+The new RMSE for linear regression 2 is 3.81 lower than 6.29 on training data, which is a sign of good prediction.
+Besides, the new RMSE of LASSO model is 6.47 which is slightly higher than training.
+Discriminant analysis has the same accuracy for both training and testing data.
 Based on all the result of the machine learning I have used on dataset, I think linear regression model with LASSO regularization is a good choice among them.
+Also, discriminant analysis has good performance on predicting in testing data.
 
@@
 I had some questions to explore and did data analysis on this dataset.
+Samples were collected randomly, and some of the data was synthetic.
+The result of data analysis indicate that both environmental factors and personal lifestyle habits can affect the degree of obesity according to bmi as a measurement.
+It could provide some suggestions from this project that, although BMI might change with aging regard to gender, if someone can do more physical exercise, drink more water and consume less alcohol, it will reduce the risk for getting obesity.
+And relevant departments and organizations can also provide corresponding recommendations and appropriate publicity to increase the awareness of obesity risk factors among residents of these countries and help them develop a healthy lifestyle.
 
@@
 
-For machine learning model fitting, I found the best predictors for BMI index based on the value of some performance metrics after trying both linear regression and classification model, which were gender, age, family history, water drinking amount, frequency of physical activity, alcohol consumption frequency and commute transporation.
+For machine learning model fitting, I found the best predictors for BMI index based on the value of some performance metrics after trying both linear regression and classification models, which were gender, age, family history, water drinking amount, frequency of physical activity, alcohol consumption frequency and commute transportation, and all of the categories were used in the analysis.
 Although the machine learning models that I used for fitting the dataset are different types and their metrics and not be compared directly, the assessment of model quality using testing data could give a clue for the model selection.
@@
 
+BMI is a fast, low cost and easy-to-operate measurement for obesity rated in a population.
+It applies to most of the people and could fit large population groups better than measuring body fat content.
 I think my analysis has include two different types of model according to the type of outcome and each type include 2 models, which can provide more information and the result can be shown in different aspects as well as more choices were provided.
@@
 
+First, there is a problem with my outcome, BMI, which [could be misleading when being used as an indicator for obesity](https://www.yalemedicine.org/news/why-you-shouldnt-rely-on-bmi-alone), because BMI value could be misleading because a person with lots of muscle and minimal body fat can have the same BMI as a person with obesity who has much less muscle.
+The other reason is that BMI was invented based on European white men and didn’t take into account that a person’s body fat also tends to vary depending on their sex, race, and ethnicity.
+So it will be better to take fat content into consideration when doing analysis for obesity.
+But it is not included.
+The difficulty for obtaining fat content might be that it consumes time to measure with specific machine.
 I did tried to do tuning for random forest model but I failed for running it successfully, so this could be a limitation that I have.

### 2024-05-03 Rcode/mlmodel-code/ml-mode.R

@@ -7,17 +7,34 @@ library(ggplot2)
 
 ## ---- loaddata --------
 #Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("starter-analysis-exercise","data","processed-data","processeddata.rds")
+data_location <- here::here("data","processed-data","processeddata.rds")
 #load data
 mydata <- readRDS(data_location)
 
 ## ---- logistic --------
-# Fit multinomial logistic regression model
-#library(nnet)
-#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)
+# Define a function to map transportation modes to numeric values
+map_transport <- function(transport_mode) {
+  if (transport_mode == "Walking") {
+    return(4)
+  } else if (transport_mode == "Motorbike") {
+    return(2)
+  } else if (transport_mode == "Public_Transportation") {
+    return(3)
+  } else if (transport_mode == "Automobile") {
+    return(1)
+  } else {
+    return(0)  # Return 0 for other cases
+  }
+}
 
-# Summary of the model
-#summary(multinom_model)
+# Apply the function to create the new variable
+mydata$MTRANSn <- sapply(mydata$MTRANS, map_transport)
+
+# Calculate correlation coefficient
+correlation <- cor(mydata$FAF, mydata$MTRANSn)
+
+# Print correlation coefficient
+print(correlation)
 
 
 ## ---- lm1 --------
@@ -60,6 +77,8 @@ saveRDS(lm4_result, file = lm_file)
 
 
 ## ---- fitting linear models --------
+# Filter out observations where Alcohol is 4
+mydata1 <- mydata[mydata$Alcohol != 4, ]
 # Split data into training and testing sets
 set.seed(123)  # Set seed for reproducibility
 trainIndex <- createDataPartition(mydata$BMI, p = 0.8, list = FALSE)

### 2024-05-04 Rcode/mlmodel-code/ml-model-v1.qmd
some minor errors

@@
 set.seed(123)
+
+# Initialize variables to store results
+minRMSE1 <- Inf
+bestSubset1 <- NULL
 
@@
 +-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
-| Linear regression 1               | Age, History, Alcohol, MTRANS                                 | RMSE=0.62     | RMSE=6.54                    |
+| Linear regression 1               | Gender, Age, History, Water, Alcohol, FAF, MTRANS             | RMSE=6.42     | RMSE=6.54                    |
 +-----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
@@
 
-# Update the formula to include Age, History, Alcohol, and MTRANS as predictors
-formula_lm1_test <- paste("BMI ~ Age + History + Alcohol + MTRANS")
+# Update the formula
+formula_lm1_test <- paste("BMI ~ Gender + Age + History + Water + Alcohol + FAF + MTRANS")
 
@@
 ```
+
 I will explore more information for LASSO model.
+
 ```{r}

### 2024-05-04 products/manuscript/Manuscript copy.qmd
some minor errors

@@
 +----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
-| Linear regression 1              | Age, History, Alcohol, MTRANS                                 | RMSE=0.62     | RMSE=6.54                    |
+| Linear regression 1              | Gender, Age, History, Water, Alcohol, FAF, MTRANS             | RMSE=6.42     | RMSE=6.54                    |
 +----------------------------------+---------------------------------------------------------------+---------------+------------------------------+
@@
 
+A former study did research on a similar topic with this project, and the analysis indicated that among men, the proportional odds of obesity increase with urban residency, aging, marital status different from single and decrease with current smoking.
+Among women, the proportional odds increase with urban residency, primary educational level, high total blood cholesterol level and high fasting blood glucose level, and decrease with current smoking [@kaboré2020].
+It proved gender, age are risk or caustic factors for overweight or obesity, as well as other environmental factors and body conditions.
+
 I found simple linear model fitting can answer my questions, which were that alcohol consumption was correlated to obesity, as well as there were interactions between physical activity frequency and water drinking amount.
@@
 Also, regularization was applied by LASSO model to reduce overfitting at the same time.
+
+For more detailed information of LASSO model, I took the review advice and tried to extract strongest and weakest coefficient, but the result showed that intercept was both strongest and weakest coefficient.
+I didn't figure out how it could be interpret so it is not included in this part.
 
@@
 The difficulty for obtaining fat content might be that it consumes time to measure with specific machine.
-I did tried to do tuning for random forest model but I failed for running it successfully, so this could be a limitation that I have.
+Doing tuning for random forest model has been tried but I failed for running it successfully, so this could be a limitation.
 And maybe there could be more metrics that I can also calculate but I only chose 1 for each model.
-Furthermore, chatGPT told me that even if testing result has a sign for well predicting, there still could be other factors for further assessment such as F1-score and so on.
+Furthermore, chatGPT said that even if testing result has a sign for well predicting, there still could be other factors for further assessment such as F1-score and so on.
 But base on my understanding, I don't know what other methods I could choose from so I just stopped here.

### 2024-05-04 products/manuscript/supplement/Supplementary-Material.qmd
some minor errors

@@
 
--   assets: storing reference pdf files and citation style files
-
+-   assets: storing reference pdf files and citation style file
 -   data:
-
     -   raw-data: codebook created by me and original data file downloaded from website
-
 -   processed-data: cleaned data saved in RDS file
-
 -   products: manuscript in qmd file and its render output with same name in doc file
-
--   R: including all codes needed
-
+-   Rcode: including all codes needed
     -   analysis-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file
-
-    -   eda-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file
-
-    -   processing-code: including codes for data cleaning process both in qmd and R script file
-
--   renv: storage for all of the packages used in this project in R
-
+    -   dneda-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file
+    -   mlmmodel-code: including codes for data cleaning process both in qmd and R script file
 -   results:
-
     -   figures: all of the plots in png file created and saved from eda-code and analysis-code
-
     -   tables: summary tables created and saved from eda-code and 2 linear model tables from analysis-code

### 2024-05-04 Rcode/mlmodel-code/ml-mode.R

@@ -228,29 +228,104 @@ for (i in 1:length(predictors_updated)) {
 print(paste("Best subset of predictors (including interactions) using cross-validation:", paste(bestSubset2, collapse = ", ")))
 print(paste("Lowest RMSE (including interactions) using cross-validation:", minRMSE2))
 
+# Create dummy variables for categorical variables in trainData
+trainData1 <- within(trainData, {
+  History <- as.factor(History)
+  MTRANS <- as.factor(MTRANS)
+})
+dummy_vars_train <- model.matrix(~ History + MTRANS - 1, data = trainData1)
+
+# Combine dummy variables with original trainData
+trainData1 <- cbind(trainData, dummy_vars_train)
+
+# Repeat the same process for testData
+testData1 <- within(testData, {
+  History <- as.factor(History)
+  MTRANS <- as.factor(MTRANS)
+})
+dummy_vars_test <- model.matrix(~ History + MTRANS - 1, data = testData1)
+testData1 <- cbind(testData, dummy_vars_test)
+
+# Define predictors and outcome
+response <- "BMI"
+predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
+interaction_terms <- c("Gender_Age", "FAF_Water")
+
+# Update predictors to include dummy variables
+predictors_lasso <- c("Gender", "Age", "Water", "Alcohol", "FAF","Gender_Age", "FAF_Water", colnames(dummy_vars_train))
+
+
+# Define function to calculate RMSE
+calculate_rmse <- function(predictions, actual) {
+  return(sqrt(mean((actual - predictions)^2)))
+}
+
 # Initialize variables to store results
 min_rmse_lasso <- Inf
 best_subset_lasso <- NULL
 
+# Set seed
+set.seed(123)
+
 # Perform LASSO regression with cross-validation
-lasso_model <- cv.glmnet(as.matrix(trainData[, predictors_updated]), trainData$BMI, alpha = 1, nfolds = 10)
+for (fold in 1:10) {
+  # Extract indices for training and validation data for current fold
+  train_indices <- cv_folds[[fold]]
+  validation_indices <- setdiff(1:nrow(trainData1), train_indices)
+  
+  # Define training and validation data subsets
+  train_data_fold <- trainData1[train_indices, ]
+  validation_data_fold <- trainData1[validation_indices, ]
+  
[…]
